### 2014-08-20 paper.tex
Initial structure of the paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%2345678901234567890123456789012345678901234567890123456789012345678901234567890
%        1         2         3         4         5         6         7         8
\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper

\IEEEoverridecommandlockouts                              % This command is only needed if 
                                                          % you want to use the \thanks command

\overrideIEEEmargins                                      % Needed to meet printer requirements.

% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

\usepackage[utf8x]{inputenc}

\usepackage[protrusion=true,expansion=true]{microtype}
\usepackage{cite}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{url}

\usepackage{minted}
\renewcommand{\theFancyVerbLine}{
  \sffamily\textcolor[rgb]{0.5,0.5,0.5}{\scriptsize\arabic{FancyVerbLine}}}

\graphicspath{{figs/}}
\DeclareGraphicsExtensions{.pdf,.jpg,.png}

\usepackage{fixme}

\newcommand{\ie}{{\textit{i.e.\ }}}
\newcommand{\cf}{{\textit{cf\ }}}
\newcommand{\eg}{{\textit{e.g.\ }}}
\newcommand{\pyRobots}{{{\sc pyRobots}}}

\title{\LARGE \bf
    {\sc pyRobots}, a executive environment for robot control
}

\author{Séverin Lemaignan, Anahita Hosseini and Pierre Dillenbourg\\
Computer-Human Interaction in Learning and Instruction \\
Ecole Polytechnique Fédérale de Lausanne (EPFL) \\
CH-1015 Lausanne, Switzerland \\
{\tt\small firstname.last@epfl.ch}
}

\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

    Blabla

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

A set of software helpers to write high-level robot controllers.

\emph{Not} a middleware: it connects to one or several middleware to actually
perform actions.

The toolkit does not aim at formal verifiability: no contribution in this field.

\subsection{Related Work: Robot Execution Control}

\begin{itemize}
    \item FSM -> SMACH
    \item agent-based approaches -> ROAR
    \item logic languages -> OpenPRS, CRAM
    \item dedicated control languages -> URBI
    \item similar attempts: teer (co-routine -> cooperative multi-threading)
    \item and the vast majority: ad-hoc scripts
\end{itemize}

\subsection{Guiding principles}

\begin{itemize}
    \item Use a de-facto standard language (Python) to ease adoption (familiar to
        many researchers, large software ecosystem, excellent integration with
        most robotic softwares, in particular common middlewares like ROS or
        YARP)
    \item should be as lightweight/transparent as possible (in particular from a
        syntax perspective) so that the
        programmer can focus on the programming of the behaviours (I'm
        watching you, SMACH)
    \item should enforce as few design choices as possible. 'convention over
        configuration' as often as possible
\end{itemize}

\subsection{Main features}

Parallelism (all \emph{actions} are asynchronous) and event-based programming
are at the core.

It also offers support for robot resources management (locking of ``parts'' of
the robot) and uniform manipulation of 6D poses (that integrates transparently
with other tools like ROS's TF when available).

\section{Overview of the framework}

\begin{itemize}
    \item Turns any Python function into a background action with the decorator
        @action.
    \item Robot actions are non-blocking by default: they are instanciated as
        futures (lightweight threads),
    \item Actions can be cancelled at any time via signals (the
        ActionCancelled signal is raised).
    \item Lock specific resources with a simple @lock(...) in front of the
        actions. When starting, actions will wait for resources to be
        available if needed.
    \item Supports compound resources (like WHEELS == LEFTWHEEL +
        RIGHTWHEEL)
    \item Create event with robot.every(<condition>).do(<action>)
    \item Poses are managed explicitely and can easily be
        transformed from one reference frame to another one
        (integrates with ROS TF when available).
    \item Extensive logging support to debug and replay
        experiments.
\end{itemize}

\section{Example}

%\begin{listing}[ht]
\begin{minted}[frame=lines,
               linenos=true,
               gobble=4,
               fontsize=\scriptsize]{python}
    import time
    from robots import GenericRobot
    from robots.decorators import action, lock
    from robots.resources import Resource
    from robots.signals import ActionCancelled

    # create a 'lockable' resource for our robot
    WHEELS = Resource("wheels")

    class MyRobot(GenericRobot):

      def __init__(self):
        super(MyRobot, self).__init__()

        # create (and set) one element in the robot's state.
        # Here a bumper.
        self.state.my_bumper = False

        # do whatever other initialization you need :-)

      def send_goal(self, pose):
        # move your robot using your favorite middleware
        print("Starting to move towards %s" % pose)

      def stop(self):
        # stop your robot using your favorite middleware
        print("Motion stopped")

      def whatever_lowlevel_method_you_need(self):
        pass

    @lock(WHEELS)
    @action
    def move_forward(robot):
      """ We write action in a simple imperative, 
          seemingly blocking way.
      """

      # the target pose: simply x += 1.0m in the robot's 
      # frame. pyRobots will handle the frames 
      # transformations as needed.
      target = [1.0, 0., 0., "base_link"]

      try:
        robot.send_goal(target)

        while(robot.pose.distance(robot.pose.myself(), 
                                  target) > 0.1):
            # robot.sleep is exactly like time.sleep, 
            # except it lets the pyrobots signals pass 
            # through.
            robot.sleep(0.5)

        print("Motion succeeded")

      except ActionCancelled:
        # if the action is cancelled, clean up your state
        robot.stop()

    with MyRobot() as robot:

      # Turn on DEBUG logging.
      # Shortcut to configure the std Python logging system
      robot.debug()

      robot.whenever("my_bumper", value = True)
                                    .do(move_forward)

      try:
        while True:
          time.sleep(0.5)
      except KeyboardInterrupt:
        pass
\end{minted}
%\caption{Example from external file}
%\label{listing:3}
%\end{listing}

\section{Experimental Deployments}

\subsection{Platforms}

\begin{itemize}
    \item PR2, with ROS and pocolibs
    \item Nao, with ROS and NAOqi
    \item Ranger with Aseba and ROS
\end{itemize}

\subsection{Experiments and studies}

\begin{itemize}
    \item roboscopie
    \item LAAS architecture
    \item CoWriter
    \item Ranger
\end{itemize}

\paragraph{Stress-test in a nursery}

\begin{figure}
        \centering
        \includegraphics[width=0.9\columnwidth]{ranger-side}
        \caption{Infants playing freely (!) with an EPFL's \emph{Ranger} robot.
        The behaviour of the robot is controlled with \pyRobots}
        \label{expe-nursery}
\end{figure}

\section{Future Work}

\begin{itemize}
    \item control mixing policies?
    \item control over action priorities
\end{itemize}

\section*{Acknowledgment}

This work has been supported by...

\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,biblio}

\end{document}

### 2014-08-20 paper.tex
Minor: formatting, macros...

@@
 
+  \newminted{python}{frame=lines,
+                     linenos=true,
+                     gobble=4,
+                     fontsize=\scriptsize}
+
 \graphicspath{{figs/}}
@@
 
-\usepackage{fixme}
+\usepackage[draft]{fixme}
 
@@
 \title{\LARGE \bf
-    {\sc pyRobots}, a executive environment for robot control
+    \pyRobots, a toolkit for robot executive control
 }
@@
 Computer-Human Interaction in Learning and Instruction \\
-Ecole Polytechnique Fédérale de Lausanne (EPFL) \\
+École Polytechnique Fédérale de Lausanne (EPFL) \\
 CH-1015 Lausanne, Switzerland \\
@@
 
-\section{Overview of the framework}
+\section{Concepts}
+
+Before introducing and explaining step-by-step a complete example in the next
+section, it is useful to introduce the main concepts of \pyRobots.
 
@@
 %\begin{listing}[ht]
-\begin{minted}[frame=lines,
-               linenos=true,
-               gobble=4,
-               fontsize=\scriptsize]{python}
+
+\begin{pythoncode}
     import time
@@
     def move_forward(robot):
-      """ We write action in a simple imperative, 
-          seemingly blocking way.
+      """ Actions are written in a simple imperative, 
+          blocking style.
       """
@@
         pass
-\end{minted}
+\end{pythoncode}
 %\caption{Example from external file}

### 2014-08-20 paper.tex
First pass on the introduction

@@
 
-A set of software helpers to write high-level robot controllers.
+Orchestrating the activity of a robot is a central issue of robotics: \emph{what
+to do when?} Traditionally, in the famous 3-layer architectures~\fixme{ref?},
+this function is implemented in the so-called \emph{deliberative layer} where
+decisions are made based on perceptions and on the current internal state, and
+executed by sending orders to a \emph{functional layer}.
 
-\emph{Not} a middleware: it connects to one or several middleware to actually
-perform actions.
+While individual decision-making components (like task planners) are studied in
+(relative) isolation since years, the \emph{orchestration} issue, with questions
+like \textit{When to start them? Which one should be selected? How to react to a
+new situation in a timely manner?} remains difficult to address in a generic
+way.
 
-The toolkit does not aim at formal verifiability: no contribution in this field.
+Many approaches have been devised (we review them in the next section) that
+include \emph{Finite State Machines} (FSM), domain-specific languages (in
+particular, logic languages) or agent-based frameworks.
+
+These tools adopt principled approaches, often with solid theoretical
+contributions. However, it seems that none of them gained broad acceptance in
+the robotic community, and many of us resort to write ad-hoc scripts, suitable
+for a single application/demonstration/experiment and neither reliable nor
+extensible.
+
+We hypothesize that the software architecture design that these tools enforce,
+combined with their general lack of practicality (unfamiliar language; lack of
+bindings for a given robot or middleware; difficulty to install or setup;
+non-trivial deployment) explain this low level of acceptance more than any
+particular intrinsic weaknesses.
+
+This article introduces our attempt at designing an unobtrusive execution control
+toolkit, that aims at addressing the \emph{practicality} issue: it has been
+designed and implemented from the bottom-up, starting from actual needs when
+running complex robots in largely unpredictable scenarios (typically,
+loosely constrained human-robot interactions), and in the real world.
+
+Instead of an \emph{environment} or a \emph{framework}, which would suggest
+strong design and development constraints, \pyRobots can be conceived as a set of
+powerful software helpers to write parallel, event-based, high-level robot
+controllers.
+
+It is written for and in Python, which ensure familiarity, fast development
+cycles and broad support for interfacing with existing robots and middlewares.
+This must be emphasized: \pyRobots is \emph{not} a middleware. It purposefully
+provides no mechanisms to connect to other modules or components. It instead
+\emph{uses} one (or several) middleware to actually communicate with the robot
+and perform actions.
+
+One the other hand, \pyRobots does not provide formal models or verifiability:
+in its current state, it does not attempt to contribute in this field. It
+focuses instead on providing unobtrusive tools to control robots in environments
+that require carrying out many tasks in parallel, in an unpredictable,
+event-based manner.
 
@@
 \begin{itemize}
-    \item FSM -> SMACH
-    \item agent-based approaches -> ROAR
-    \item logic languages -> OpenPRS, CRAM
-    \item dedicated control languages -> URBI
-    \item similar attempts: teer (co-routine -> cooperative multi-threading)
+    \item FSM: SMACH
+    \item agent-based approaches: ROAR
+    \item logic languages: OpenPRS, CRAM
+    \item dedicated control languages: URBI
+    \item similar attempts: teer (co-routine: cooperative multi-threading)
     \item and the vast majority: ad-hoc scripts

### 2014-08-20 paper.tex
Added minted.sty 2.0alpha to have inline python code

@@
                     fontsize=\scriptsize}
+
+\newmintinline[python]{python}{fontsize=\footnotesize}
 
@@
     \item Turns any Python function into a background action with the decorator
-        @action.
+        \python|@action|.
     \item Robot actions are non-blocking by default: they are instanciated as
@@
     \item Actions can be cancelled at any time via signals (the
-        ActionCancelled signal is raised).
-    \item Lock specific resources with a simple @lock(...) in front of the
+        \python|ActionCancelled| signal is raised).
+    \item Lock specific resources with a simple \python|@lock(...)| in front of the
         actions. When starting, actions will wait for resources to be
         available if needed.
-    \item Supports compound resources (like WHEELS == LEFTWHEEL +
-        RIGHTWHEEL)
+    \item Supports compound resources (like \python|WHEELS == LEFTWHEEL + RIGHTWHEEL|)
     \item Create event with robot.every(<condition>).do(<action>)

### 2014-08-20 paper.tex
Minor, including fig caption

@@
 
-\usepackage{minted}
+\usepackage[cache]{minted}
 \renewcommand{\theFancyVerbLine}{
@@
                     gobble=4,
-                    fontsize=\scriptsize}
+                    fontsize=\scriptsize,
+                    xleftmargin=1.8em}
 
@@
 \newcommand{\eg}{{\textit{e.g.\ }}}
-\newcommand{\pyRobots}{{{\sc pyRobots}}}
+\newcommand{\pyRobots}{\textsc{pyRobots}\ }
 
@@
         \caption{Infants playing freely (!) with an EPFL's \emph{Ranger} robot.
-        The behaviour of the robot is controlled with \pyRobots}
+        The behaviour of the robot is controlled with \pyRobots. It sustained an
+        average of 70 actions triggered by minute over more than an hour,
+        with peaks at 200 actions/minute.}
         \label{expe-nursery}

### 2014-08-20 paper.tex
WIP on the Concepts and Overview

@@
 
-\section{Concepts}
+\section{Concepts and Overview}
 
-Before introducing and explaining step-by-step a complete example in the next
-section, it is useful to introduce the main concepts of \pyRobots.
+Before introducing and explaining step-by-step a complete working example in the next
+section, it is useful to first introduce the main concepts of \pyRobots, and
+provide an overview of the features and mechanisms.
 
@@
     \item Create event with robot.every(<condition>).do(<action>)
-    \item Poses are managed explicitely and can easily be
+    \item Poses are managed explicitly and can easily be
         transformed from one reference frame to another one
@@
 
-\section{Example}
+\subsection{Core concepts}
+
+\pyRobots core concepts are simple: on one hand, the user creates a
+\textbf{robot} as an instance of a \python|GenericRobot|, that encapsulates the
+different low-level \textbf{controllers} (proxies to actuators, typically
+provided by a middleware like ROS) and the \textbf{state} of the robot
+(typically a set proxies to the sensors, or a connector to a database, etc.).
+
+On the other hand, the user creates as many high-level \textbf{actions} as
+desired, as regular Python functions. Actions are automatically added to the
+\textbf{robot}, and can access its state and controllers to perform actual
+physical actions.
+
+The user can list \textbf{events} that the \textbf{robot} must monitor, and
+attach callback to them.
+
+Finally, the user can declare arbitrary \textbf{resources}, usually
+corresponding to the robot components (for instance, the wheels, the camera,...)
+and guarantee exclusive access to the required resource to specific actions.
+
+\subsection{Asynchronous actions}
+
+As they will be executed in independent
+threads, these functions may block or never return (typically useful to
+implement lasting background behaviours).
+
+Actions can also call each-other, allowing decomposition of
+complex tasks into sub-actions and eventually atomic actions.
+
+\subsection{Events}
+
+\subsection{Resources management}
+
+\subsection{Uniform management of rigid transformations and frames}
+
+\subsection{Developer support: logging, debugging, introspection}
+\label{}
+
+\section{Implementation and Example}
 

### 2014-08-20 paper.tex
WIP async actions

@@
     \item Supports compound resources (like \python|WHEELS == LEFTWHEEL + RIGHTWHEEL|)
-    \item Create event with robot.every(<condition>).do(<action>)
+    \item Create event with \python|robot.every(<condition>).do(<action>)|
     \item Poses are managed explicitly and can easily be
@@
 
-As they will be executed in independent
-threads, these functions may block or never return (typically useful to
-implement lasting background behaviours).
+\emph{Actions} in \pyRobots are always asynchronous. They are implemented as
+\emph{futures} (also known as \emph{promises}) and therefore executed in
+independent threads. To the developer, however, they appear as regular Python
+function, simply annotated with the \python|@action| decorator, and are invoked
+as any other function.  Since they are asynchronous, these functions may
+perfectly block or even never return (typically useful to implement lasting
+background behaviours).
 
-Actions can also call each-other, allowing decomposition of
-complex tasks into sub-actions and eventually atomic actions.
+Actions can also call each-other, allowing natural decomposition of complex
+tasks into sub-actions and eventually atomic actions.
+
+\paragraph{Signaling} Running behaviours often need to be paused of cancelled.
+This poses a particular challenge in robotics, because before suspending an
+action, one often want to bring back the robot in some form of rest state (you
+would not want to suddenly cancel a \python|walk| action in the middle of a step
+without first bringing back the foot to the ground).
+
+\pyRobots addresses this issue by the mean of signals, and extends standard
+Python thread into \emph{signaling threads}. Listing~\ref{lst:signals}
+illustrates this mechanism.
+
+\begin{listing}[H]
+\begin{pythoncode}
+    @action
+    def my_action(robot):
+      try:
+        robot.do_something_dangerous()
+      except ActionCancelled:
+        robot.go_back_to_rest_pose()
+
+    action = robot.my_action()
+    time.sleep(1)
+    action.cancel()
+\end{pythoncode}
+\caption{Handling a cancellation signal}
+\label{lst:signals}
+\end{listing}
 

### 2014-08-21 paper.tex
Work on the 'Experimental Deployments'

@@
 
-Before introducing and explaining step-by-step a complete working example in the next
-section, it is useful to first introduce the main concepts of \pyRobots, and
-provide an overview of the features and mechanisms.
+Before introducing and explaining step-by-step a complete working example in the
+next section, it is useful to first introduce the main concepts of \pyRobots,
+and provide an overview of the features and mechanisms. Since \pyRobots is an
+open-source software, the interested reader can also refer to the source code,
+available here:~\url{https://github.com/chili-epfl/pyrobots}.
 
@@
 
-\subsection{Platforms}
+The development of \pyRobots started in 2011, driven by the requirements of a
+18-min long theatre performance between a human actor and a robot. Since that time, the toolkit has
+iteratively matured, and has been deployed and used on several projects and
+platforms. We first briefly recap these deployments with references to the
+corresponding publications, and then present our latest study, in a nursery,
+that recently acted as a stress-test for the library.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{table}[ht!]
+\begin{center}
+\begin{tabular}{p{0.9\columnwidth}}
+\hline
+    {\bf Manipulation} \\
+     {\tt attachobject}, {\tt basicgive}, {\tt basictake}, {\tt close\_gripper}, {\tt configure\_grippers}, {\tt grab\_gripper}, {\tt handover}, {\tt hide}, {\tt open\_gripper}, {\tt pick}, {\tt put}, {\tt put\_accessible}, {\tt release\_gripper}, {\tt show}, {\tt take} \\
+\hline
+    {\bf Gaze control} \\
+     {\tt glance\_to}, {\tt look\_at}, {\tt sweep}, {\tt switch\_active\_stereo\_pair}, {\tt track}, {\tt cancel\_track} \\
+\hline
+    {\bf Navigation} \\
+     {\tt carry}, {\tt follow}, {\tt cancel\_follow}, {\tt goto}, {\tt moveclose}, {\tt waypoints} \\
+\hline
+    {\bf Local navigation} \\
+     {\tt dock}, {\tt rotate}, {\tt translate} \\
+\hline
+    {\bf Posture configuration} \\
+     {\tt extractpose}, {\tt idle}, {\tt manipose}, {\tt movearm}, {\tt rest}, {\tt setpose}, {\tt settorso}, {\tt tuckedpose} \\
+\hline
+\end{tabular}
+\end{center}
+\caption{Examples of \pyRobots actions, mostly stemming from experiments involving
+    the PR2 robot.}
+
+\label{pyrobots_actions}
+\end{table}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\paragraph{Roboscopie} The first complex deployment scenario for \pyRobots was indeed a theatre
+performance entitled \emph{Roboscopie}~\cite{lemaignan2012roboscopie}. The
+toolkit was created out of the need to control a complex robot (a Willow Garage
+PR2) in a structured way, as required for a live performance with a
+human\footnote{The original \pyRobots script used for the performance can be
+accessed from~\url{http://www.laas.fr/roboscopie}. Note that this script does
+not reflect anymore the toolkit usage.}
+
+A total of 47 \pyRobots actions were written for this event
+(Table~\ref{pyrobots_actions} gives a few examples), and the toolkit proved to
+be especially convenient to interleave calls to both ROS-based and {\sc
+GenoM}-based\footnote{{\sc GenoM} is software component as well as a middleware
+developed at LAAS-CNRS} components through their respective Python bindings.
+
+\pyRobots also allowed us to retain the \emph{rapid prototyping} convenience of a
+script language, essential for the robot to ``practise'' and adapt to the
+vision of the director.
 
@@
 
-\subsection{Experiments and studies}
-
 \begin{itemize}
@@
 \end{itemize}
+
+\begin{figure}
+        \centering
+        \includegraphics[width=0.9\columnwidth]{cowriter}
+        \caption{Children teaching An Aldebaran Nao interacting with }
+        \label{expe-cowriter}
+\end{figure}
 

### 2014-08-22 paper.tex
Work on events, resource and poses sections

@@
 \usepackage{url}
+\usepackage{amsmath}
 
@@
 \begin{itemize}
-    \item Turns any Python function into a background action with the decorator
-        \python|@action|.
-    \item Robot actions are non-blocking by default: they are instanciated as
-        futures (lightweight threads),
-    \item Actions can be cancelled at any time via signals (the
-        \python|ActionCancelled| signal is raised).
     \item Lock specific resources with a simple \python|@lock(...)| in front of the
@@
     \item Supports compound resources (like \python|WHEELS == LEFTWHEEL + RIGHTWHEEL|)
-    \item Create event with \python|robot.every(<condition>).do(<action>)|
     \item Poses are managed explicitly and can easily be
@@
 provided by a middleware like ROS) and the \textbf{state} of the robot
-(typically a set proxies to the sensors, or a connector to a database, etc.).
+(typically a set proxies to the sensors, or a connector to a database, etc.,
+accessed through a dictionary-like object).
 
@@
 function, simply annotated with the \python|@action| decorator, and are invoked
-as any other function.  Since they are asynchronous, these functions may
+like any other function.  Since they are asynchronous, these functions may
 perfectly block or even never return (typically useful to implement lasting
@@
 
-Actions can also call each-other, allowing natural decomposition of complex
-tasks into sub-actions and eventually atomic actions.
+Actions can also call each other, allowing natural decomposition of complex
+tasks into sub-actions and atomic actions.
 
-\paragraph{Signaling} Running behaviours often need to be paused of cancelled.
-This poses a particular challenge in robotics, because before suspending an
-action, one often want to bring back the robot in some form of rest state (you
-would not want to suddenly cancel a \python|walk| action in the middle of a step
-without first bringing back the foot to the ground).
+\paragraph{Signaling} Behaviours often need to be paused of cancelled.  This
+raises an issue that needs to be addressed in robotics: before suspending an
+action, one often wants to first bring back the robot into some form of rest
+state (\eg one would not want to suddenly cancel a \python|walk| action in the
+middle of a step without first bringing back the foot to the ground).
 
-\pyRobots addresses this issue by the mean of signals, and extends standard
-Python thread into \emph{signaling threads}. Listing~\ref{lst:signals}
-illustrates this mechanism.
+\pyRobots addresses this issue through signals, and extends standard
+Python thread into \emph{signaling threads}, relying on Python's usual syntax
+for exception handling. Listing~\ref{lst:signals}
+illustrates this mechanism. The action is instantiated and started at line 8,
+and cancelled at line 10. This send a \python|ActionCancelled| signal to the
+action that the user can handle using an \python|except| statement.
 
@@
       try:
-        robot.do_something_dangerous()
+        robot.do_something_long()
       except ActionCancelled:
@@
 
+Pausing actions is more complex (the action internal state must be saved, a mean
+to resume the action where it was interrupted must be provided, the user must be
+given ways to code transition behaviours when pausing and resuming) and is not
+currently supported by \pyRobots.
+
 \subsection{Events}
+
+Sharing the same syntax with URBI, events in \pyRobots can by monitored through
+the methods \python|on| (one shot event) and \python|whenever| (continuous event
+monitoring). Event conditions can be either a predicate (that must take as
+single argument the robot instance) or a pair of $\{\text{state member},
+\text{condition}\}$ with condition in $\{\text{value}=x, \text{become}=x,
+\text{below}=x, \text{above}=x, \text{decrease}=x, \text{increase}=x\}$. An
+example of event usage can be seen at line 67 of listing~\ref{mwe}.
+
+When the event condition is verified, the callback (usually, an action) is
+invoked from the event monitor thread.
 
@@
 
-\subsection{Uniform management of rigid transformations and frames}
+Usually, physical robot resources (mainly actuators, but possibly sensors as
+well) require exclusive use (it is seldom desirable that two different actions
+control the wheels of a robot at the same time).
+
+To leave full control regarding the required resource modeling granularity, the
+programmer define them himself, either as stand-alone resources or as compound
+resources (listing~\ref{lst:resources}).
+
+\begin{listing}[H]
+\begin{pythoncode}
+    L_ARM = Resource()
+    R_ARM = Resource()
+    ARMS = CompoundResource(L_ARM, R_ARM)
+
+    @action
+    @lock(ARMS)
+    def lift_box(robot):
+        #...
+
+    @action
+    @lock(L_ARM)
+    def wave_hand(robot):
+        #...
+
+    @action
+    @lock(L_ARM, wait=False)
+    def scratch_head(robot):
+        #...
+
+    robot.lift_box()
+    robot.wave_hand() # waits until lift_box is over
+    robot.scratch_head() # skipped
+
+\end{pythoncode}
+\caption{Resource usage is defined at the action-level, through annotations.}
+\label{lst:resources}
+\end{listing}
+
+The user then annotates the actions with the (possibly multiple) resources they
+require (lines 6, 11 and 16 of listing~\ref{lst:resources}). When executed, the
+action acquires a lock on the resource (and all dependent resources in the case
+of a compound resource) which is released upon action completion. In the
+meantime, other actions requiring the same resource will either wait until the
+resource is release (line 21) or be skipped (line 22) if the \python|@lock|
+annotation specifies not to wait (line 16).
+
+\subsection{Uniform management of poses and rigid transformations}
+
+Managing poses, transformations, frames in real-world scenarios is not an easy
+task, mainly because it seems that every single library, middleware or robotic
+platform rely on its own conventions (rotation representations, in particular,
+are a constant source of frustration, having to jump all the time between
+quaternions, one of the 23 Euler conventions, rotation matrices, Rodrigues,
+yaw-pitch-roll notation...). Within the ROS ecosystem, the Transform Library
+(tf,~\cite{foote2013tf}) has played a major role to make it convenient to work
+with frames, and poses can always be expressed and shared in the most convenient
+local frame, tf taking care of required intermediate transformation.
+
+\pyRobots offers a similar approach \emph{within} the control script, adding a
+pinch of syntactic sugar to make poses and frame management transparent for the
+high-level action controller.
+
+As illustrated in listing~\ref{lst:poses}, poses can be flexibly defined by a
+simple frame name, by a triple $\{x, y, z\}$ (default frame {\tt map} is then
+assumed), by a 4-tuple $\{x, y, z, frame\}$, by a 6-tuple $\{x, y, z, rx, ry,
+rz\}$ (using $sxyz$ Euler convention), by a 7-tuple $\{x, y, z, rx, ry, rz,
+frame\}$, by a 7-tuple $\{x, y, z, qx, qy, qz, qw\}$, by a 8-tuple $\{x, y, z,
+qx, qy, qz, qw, frame\}$ or by a dictionary that defines any of these 8 values.
+Internally, \pyRobots normalizes and stores the poses in their full 8-tuple
+expansion.
+
+\begin{listing}[H]
+    \begin{pythoncode*}{linenos=false, xleftmargin=0em}
+
+    # {'x':0., 'y':0., 'z':0., 
+    #  'qx':0., 'qy':0., 'qz':0., 'qw':1., 
+    #  'frame': 'handL'}
+    robot.poses["handL"]
+
+    # {'x':1., 'y':0., 'z':0., 
+    #  'qx':0., 'qy':0., 'qz':0., 'qw':1., 
+    #  'frame': 'map'}
+    robot.poses[(1.0, 0.0, 0.0)]
+
+    # {'x':0., 'y':0., 'z':1., 
+    #  'qx':0., 'qy':0., 'qz':0.0998, 'qw':0.0995, 
+    #  'frame': 'handL'}
+    robot.poses[{'z':1.0, 'rz':0.2, 'frame':'handL'}]
+
+\end{pythoncode*}
+\caption{Examples of poses normalizing.}
+\label{lst:poses}
+\end{listing}
+
+Using helper methods like \python|robot.poses.inframe(<pose>, <other frame>)| or
+\python|robot.poses.pantilt(<pose>, <ref frame>)|, the programmer can then
+transform a given pose into a different frame or different representation.
+
+These transformations are directly computed, but \pyRobots also takes advantage
+of middleware-specific infrastructure (when available) to provide known frames.
+ROS TF, Aldebaran's {\tt NAOqi} frame conventions, etc. are used when available
+through specializations of the \pyRobots \python|PoseManager| class.
 
@@
 
-\section{Implementation and Example}
+\section{Detailed Example}
 
-%\begin{listing}[ht]
+\begin{listing}[h!]
 
@@
         # create (and set) one element in the robot's state.
-        # Here a bumper.
-        self.state.my_bumper = False
+        # Here a sonar.
+        self.state["sonar"] = 10.
 
@@
 
-      robot.whenever("my_bumper", value = True)
-                                    .do(move_forward)
+      robot.whenever("sonar", below = 0.3).do(move_forward)
 
@@
 \end{pythoncode}
-%\caption{Example from external file}
-%\label{listing:3}
-%\end{listing}
+\caption{Full example of a \pyRobots application.}
+\label{mwe}
+\end{listing}
 

### 2014-09-02 paper.tex
WIP state of the art

@@
 
-\begin{itemize}
-    \item FSM: SMACH
-    \item agent-based approaches: ROAR
-    \item logic languages: OpenPRS, CRAM
-    \item dedicated control languages: URBI
-    \item similar attempts: teer (co-routine: cooperative multi-threading)
-    \item and the vast majority: ad-hoc scripts
-\end{itemize}
+Any robotic experiment requires a certain amount of high-level control to
+supervise the robot behaviour, and, expectedly, many projects have attempted to
+design and provide tools for this task.  We focus this review on \emph{tools
+dedicated to the explicit implementation of high-level behaviours}, excluding
+complete control stacks (that would typically include a middleware and hardware
+abstractions), meta-tools (like component generators or model checkers) and
+(semi-)automatic techniques (like learning by demonstration).  In contrast with
+middlewares for which the robotic community relies on a few \emph{de-facto}
+standards (ROS, YARP, OpenRTM), we want to underline that no such widely-used
+softwares exist for execution control, and we believe that most of today's robot
+behaviours are implemented as \emph{ad-hoc} scripts.
+
+We identify three categories of execution control tools: \emph{finite state
+machines}, \emph{agent-based} controllers and
+\emph{domain-specific languages} (DSL), including \emph{embedded domain-specific
+languages} (eDSL,~\cite{joyeux2011robot}). These categories overlap to a certain
+extend, but they should reasonably reflect the motivational concepts/ideas that
+gave birth to the specific tools.
+
+\paragraph{Finite State Machines} (or FSM) are the most well-known paradigm to
+describe how a robot goes from one \emph{state} (typically, performing an
+action) to another state when some condition becomes true or some event is
+triggered. Common FSM implementations in robotics include {\sc
+rfSM}~\cite{klotzbucher2010orocos} (part of the OROCOS framework) and
+{\sc smach}~\cite{bohren2010smach}. State machine semantics are well
+understood~\cite{harel1996statemate} and provide a well structured model,
+suitable for meta-reasoning (like verification). However, FSM are also
+acknowledged to be ill-suited for complex systems (where the number of state
+easily ``explodes'') or relatively dynamic and unpredictable environments where
+it can prove difficult to \textit{a priori} exhaustively list transitions.
+
+\paragraph{Agent-based controllers}
+
+temporal planning/temporal constraints: T-REX~\cite{mcgann2007trex} -> sensing-planning-acting
+
+ROAR: ~\cite{degroote2011roar}
+
+\paragraph{Domain-specific languages}
+
+Domain-specific languages (DSL), in the widest sense, represent the largest body
+of literature on high-level behaviour programming. They can either be ``actual''
+domain-specific languages like PRS~\cite{ingrand1996prs},
+RPL~\cite{mcdermott1993reactive}, Golog~\cite{levesque1997golog} (and its
+derivative like {\sc ReadyLOG}~\cite{ferrein2007robot} or, more recently,
+URBI~\cite{baillie2005urbi}; or what Joyeux calls \emph{embedded}
+DSL~\cite{joyeux2011robot}: extensions of existing languages (typically, dynamic
+languages like Python or Ruby, or logic languages like Prolog) to provide
+high-level paradigms and control structures well suited to the programming of
+robots. {\sc cram}~\cite{beetz2010cram} (based on Lisp and Prolog), {\sc
+roby}~\cite{joyeux2009plan} (based on Ruby), {\sc
+teer}~\footnote{\url{http://wiki.ros.org/executive_teer}} (based on Python), are
+examples of such eDSL.
+
+Projects like PRS, Golog or {\sc cram} have large scopes:
+
+\pyRobots positions itself in this landscape as a lightweight eDSL (in a similar
+fashion to {\sc roby} or {\sc teer}), that focuses on high-level behaviours
+implementation. It seeks unobstrusive interaction with the other components of
+the robot software stack, using Python as ubiquitous binding language. Similar
+to URBI, it focuses on transparent preemptive concurrency and event-based programming.
+Also, \pyRobots' model favours hierarchical task composition over plans, and does not
+provide by itself any task planning service, delegating this to external tools.
 
@@
     robot.wave_hand() # waits until lift_box is over
-    robot.scratch_head() # skipped
+    robot.scratch_head() # skipped if lift_box 
+                         # is still running
 
@@
 
-    # {'x':0., 'y':0., 'z':0., 
+    robot.poses["handL"]
+    # -> {'x':0., 'y':0., 'z':0., 
     #    'qx':0., 'qy':0., 'qz':0., 'qw':1., 
     #    'frame': 'handL'}
-    robot.poses["handL"]
 
-    # {'x':1., 'y':0., 'z':0., 
+    robot.poses[(1.0, 0.0, 0.0)]
+    # -> {'x':1., 'y':0., 'z':0., 
     #    'qx':0., 'qy':0., 'qz':0., 'qw':1., 
     #    'frame': 'map'}
-    robot.poses[(1.0, 0.0, 0.0)]
 
-    # {'x':0., 'y':0., 'z':1., 
+    robot.poses[{'z':1.0, 'rz':0.2, 'frame':'handL'}]
+    # -> {'x':0., 'y':0., 'z':1., 
     #    'qx':0., 'qy':0., 'qz':0.0998, 'qw':0.0995, 
     #    'frame': 'handL'}
-    robot.poses[{'z':1.0, 'rz':0.2, 'frame':'handL'}]
 
@@
 \end{center}
-\caption{Examples of \pyRobots actions, mostly stemming from experiments involving
+\caption{Examples of \pyRobots actions, mostly extracted from experiments involving
     the PR2 robot.}
@@
 
-\paragraph{Roboscopie} The first complex deployment scenario for \pyRobots was indeed a theatre
-performance entitled \emph{Roboscopie}~\cite{lemaignan2012roboscopie}. The
-toolkit was created out of the need to control a complex robot (a Willow Garage
-PR2) in a structured way, as required for a live performance with a
-human\footnote{The original \pyRobots script used for the performance can be
-accessed from~\url{http://www.laas.fr/roboscopie}. Note that this script does
-not reflect anymore the toolkit usage.}
+\paragraph{Roboscopie} The first complex deployment scenario for \pyRobots was
+indeed a theatre performance entitled
+\emph{Roboscopie}~\cite{lemaignan2012roboscopie}. The toolkit was created out of
+the need to control a complex robot (a Willow Garage PR2) in a structured way,
+as required for a live performance with a human\footnote{The original \pyRobots
+script used for the performance can be accessed
+from~\url{http://www.laas.fr/roboscopie}. Note that this script does not
+reflect anymore the toolkit usage.}
 
@@
         \includegraphics[width=0.9\columnwidth]{cowriter}
-        \caption{Children teaching An Aldebaran Nao interacting with }
+        \caption{Children teaching an Aldebaran Nao interacting how to
+        handwrite.}
         \label{expe-cowriter}
@@
 
-\section{Future Work}
+\section{Conclusion and Future Work}
+
+\subsection{Summary and main contributions}
+
+\begin{itemize}
+    \item strives to simplicity (from a user perspective)
+    \item Middleware independent, real-world deployments with with ROS, GenoM, NAOqi, aseba
+\end{itemize}
+
+\subsection{Future work}
 
@@
     \item control over action priorities
+    \item path to more complex control tool (T-REX, etc)
 \end{itemize}
@@
 \bibliographystyle{IEEEtran}
-\bibliography{IEEEabrv,biblio}
+\bibliography{pyrobots}
 

### 2014-09-02 paper.tex
Added plot of croquignole expe
WIP on the Croquignole expe

@@
 
+We organized in June 2014 a pilot study in a kindergarten (on campus) to
+stress-test the hardware and software of the Ranger
+robot~\cite{mondada2014ranger}. The Ranger robot is an autonomous robot designed
+for interaction with children. As it was primarily designed to support children
+with tidying up, it is shaped as a wooden ``box on wheels''
+(Figure~\ref{expe-nursery}). Hardware-wise, it features an RGB-D camera, IR
+sensors, a range and bearing sensor, a physical front bumper, a scale, a
+removable pacifier (used as a soft on/off switch by the children) and it is
+covered with capacitive tactile panels. The wooden panels cover 186 RGB LEDs,
+enabling various light patterns, it can play sounds, the two eyes can be fully
+controlled (2 DoFs pupils and actuated eye-lids). Wheel controllers are designed
+such as they switch to freewheel as soon as the robot is pushed by the children.
+The robot is powered by an embedded Gumstix Overo (ARM7l) board running a custom
+Linux image, assisted by three custom microcontroller boards (for low-level
+hardware interfacing).
+
+Software-wise, all the user-facing behaviours are implemented with \pyRobots,
+and the communication with the hardware is abstracted through a mix of ROS
+and Aseba~\cite{magnenat2011aseba} nodes.
+
 \begin{figure}
@@
         \caption{Infants playing freely (!) with an EPFL's \emph{Ranger} robot.
-        The behaviour of the robot is controlled with \pyRobots. It sustained an
-        average of 70 actions triggered by minute over more than an hour,
-        with peaks at 200 actions/minute.}
+        The behaviour of the robot is controlled with \pyRobots.}
         \label{expe-nursery}
+\end{figure}
+
+We brought two Ranger robots to a kindergarten (17 children, 3 to 4 years old),
+and after a brief introduction (showing them that removing the pacifier would
+``wake up'' the robot, and putting it back would make it ``fall asleep''), we
+invited the children to freely play with the robots and integrate them into
+their usual games. The experiment lasted 68 minutes.
+
+We purposefully designed a simple behavioural scheme
+(Listing~\ref{croquignole_no_move}): two background actions
+(\python|background_blink| and \python|look_at_touches|) run continuously, and 4
+events are monitored (pacifier added/removed, toy added and bumper hit).
+
+\begin{listing}[h!]
+
+\begin{pythoncode}
+    @action
+    def on_pacifier(robot):
+        robot.look_at_lolette()
+        robot.blink()
+        sleep = robot.fall_asleep()
+        robot.lightbar(ramp=colors.RAINBOW).wait()
+        sleep.wait()
+
+    @action
+    def on_pacifier_removed(robot):
+        robot.light_bar(colors.from_hls(rand(0,1),0.8,1))
+        robot.wakeup().wait()
+        robot.move(0.4, v = 0.8).wait()
+        robot.idle().wait()
+
+    @action
+    def on_bumper(robot):
+        pulse = robot.pulse_row(0, (128, 0, 128))
+        while abs(robot.state.v) > 0.01:
+            robot.sleep(0.2)
+        pulse.cancel()
+
+    @action
+    def on_toy(robot):
+        robot.playsound(SOUNDS["toy_in"])
+        robot.lightbar(ramp=colors.RAINBOW).wait()
+
+    with Ranger() as robot:
+
+        robot.background_blink()
+        robot.look_at_touches()
+
+        robot.every("pacifier", becomes = True).do(on_pacifier)
+        robot.every("pacifier", becomes = False)
+                                       .do(on_pacifier_removed)
+        robot.every("scale", increase = 0.3).do(on_toy)
+        robot.every("bumper", becomes = True).do(on_bumper)
+
+        while True:
+            time.sleep(0.1)
+\end{pythoncode}
+\caption{Source of the high-level behaviours running on the robots during the
+nursery pilot (some behaviours like battery management have been omitted for
+clarity).}
+\label{croquignole_no_move}
+\end{listing}
+
+\begin{figure}
+    \resizebox{\columnwidth}{!}{\input{croquignole-expe}}
+    \caption{Events fired and actions started per minute, over the 68 min-long
+    experiment. The robot sustains an average of 70 actions triggered by minute over
+    more than an hour, with peaks at 200 actions/minute.}
+
 \end{figure}

### 2014-09-03 paper.tex
Mean and SD age of kindergarten expe

@@
 
-We brought two Ranger robots to a kindergarten (17 children, 3 to 4 years old),
-and after a brief introduction (showing them that removing the pacifier would
-``wake up'' the robot, and putting it back would make it ``fall asleep''), we
-invited the children to freely play with the robots and integrate them into
-their usual games. The experiment lasted 68 minutes.
+We brought two Ranger robots to a kindergarten (17 children, mean age: 3.88
+years old, SD=0.65), and after a brief introduction (showing them that removing
+the pacifier would ``wake up'' the robot, and putting it back would make it
+``fall asleep''), we invited the children to freely play with the robots and
+integrate them into their usual games. The experiment lasted 68 minutes.
 

### 2014-09-03 paper.tex
Section on logging/debugging

@@
 \end{pythoncode*}
-\caption{Examples of poses normalizing.}
+\caption{Examples of pose normalization.}
 \label{lst:poses}
@@
 
+Running robots in real-world, \ie dynamic and unpredictable, environments
+entails repetitive coding-testing-debugging development cycles. \pyRobots
+provides several tools to support this development phase. A special attention
+was paid to the logging infrastructure. It relies on the versatile Python
+built-in logging tools, which support configurable parallel logging streams
+(typically, a full debug stream is logged to a file while a less verbose stream
+is output on the console at run-time).
+
+\pyRobots logs are designed to allow full \emph{post-mortem} replay, and we
+provide a graphical interface to display the actions and resource usage over
+time (Figure~\ref{log_view}).
+
+Debugging and introspection support is provided at run-time through a range of
+off-the-shell Python debuggers. Remote debugging is also supported with standard
+tools like {\sc rpdb2}~\footnote{\url{http://winpdb.org}}. Thanks to the dynamic
+nature of Python, these debuggers not only provide insights on failures, but
+also provide full introspection and run-time manipulation of the application
+state.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Detailed Example}

### 2014-09-04 paper.tex
Added a screenshot of Anahita's log analyser
Commented out the full example. Not very useful in such a paper

@@
 
+\begin{figure}
+        \centering
+        \includegraphics[width=0.9\columnwidth]{log}
+        \caption{Screenshot of \pyRobots log analyser, processing the log of the
+        experiment presented at section~\ref{croq}}
+        \label{log_view}
+\end{figure}
+
 Debugging and introspection support is provided at run-time through a range of
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Detailed Example}
-
-\begin{listing}[h!]
-
-\begin{pythoncode}
-    import time
-    from robots import GenericRobot
-    from robots.decorators import action, lock
-    from robots.resources import Resource
-    from robots.signals import ActionCancelled
-
-    # create a 'lockable' resource for our robot
-    WHEELS = Resource("wheels")
-
-    class MyRobot(GenericRobot):
-
-      def __init__(self):
-        super(MyRobot, self).__init__()
-
-        # create (and set) one element in the robot's state.
-        # Here a sonar.
-        self.state["sonar"] = 10.
-
-        # do whatever other initialization you need :-)
-
-      def send_goal(self, pose):
-        # move your robot using your favorite middleware
-        print("Starting to move towards %s" % pose)
-
-      def stop(self):
-        # stop your robot using your favorite middleware
-        print("Motion stopped")
-
-      def whatever_lowlevel_method_you_need(self):
-        pass
-
-    @lock(WHEELS)
-    @action
-    def move_forward(robot):
-      """ Actions are written in a simple imperative, 
-          blocking style.
-      """
-
-      # the target pose: simply x += 1.0m in the robot's 
-      # frame. pyRobots will handle the frames 
-      # transformations as needed.
-      target = [1.0, 0., 0., "base_link"]
-
-      try:
-        robot.send_goal(target)
-
-        while(robot.pose.distance(robot.pose.myself(), 
-                                  target) > 0.1):
-            # robot.sleep is exactly like time.sleep, 
-            # except it lets the pyrobots signals pass 
-            # through.
-            robot.sleep(0.5)
-
-        print("Motion succeeded")
-
-      except ActionCancelled:
-        # if the action is cancelled, clean up your state
-        robot.stop()
-
-    with MyRobot() as robot:
-
-      # Turn on DEBUG logging.
-      # Shortcut to configure the std Python logging system
-      robot.debug()
-
-      robot.whenever("sonar", below = 0.3).do(move_forward)
-
-      try:
-        while True:
-          time.sleep(0.5)
-      except KeyboardInterrupt:
-        pass
-\end{pythoncode}
-\caption{Full example of a \pyRobots application.}
-\label{mwe}
-\end{listing}
+% \section{Detailed Example}
+% 
+% \begin{listing}[h!]
+% 
+% \begin{pythoncode}
+%     import time
+%     from robots import GenericRobot
+%     from robots.decorators import action, lock
+%     from robots.resources import Resource
+%     from robots.signals import ActionCancelled
+% 
+%     # create a 'lockable' resource for our robot
+%     WHEELS = Resource("wheels")
+% 
+%     class MyRobot(GenericRobot):
+% 
+%       def __init__(self):
+%         super(MyRobot, self).__init__()
+% 
+%         # create (and set) one element in the robot's state.
+%         # Here a sonar.
+%         self.state["sonar"] = 10.
+% 
+%         # do whatever other initialization you need :-)
+% 
+%       def send_goal(self, pose):
+%         # move your robot using your favorite middleware
+%         print("Starting to move towards %s" % pose)
+% 
+%       def stop(self):
+%         # stop your robot using your favorite middleware
+%         print("Motion stopped")
+% 
+%       def whatever_lowlevel_method_you_need(self):
+%         pass
+% 
+%     @lock(WHEELS)
+%     @action
+%     def move_forward(robot):
+%       """ Actions are written in a simple imperative, 
+%           blocking style.
+%       """
+% 
+%       # the target pose: simply x += 1.0m in the robot's 
+%       # frame. pyRobots will handle the frames 
+%       # transformations as needed.
+%       target = [1.0, 0., 0., "base_link"]
+% 
+%       try:
+%         robot.send_goal(target)
+% 
+%         while(robot.pose.distance(robot.pose.myself(), 
+%                                   target) > 0.1):
+%             # robot.sleep is exactly like time.sleep, 
+%             # except it lets the pyrobots signals pass 
+%             # through.
+%             robot.sleep(0.5)
+% 
+%         print("Motion succeeded")
+% 
+%       except ActionCancelled:
+%         # if the action is cancelled, clean up your state
+%         robot.stop()
+% 
+% 
+%     with MyRobot() as robot:
+% 
+%       # Turn on DEBUG logging.
+%       # Shortcut to configure the std Python logging system
+%       robot.debug()
+% 
+%       robot.whenever("sonar", below = 0.3).do(move_forward)
+% 
+%       try:
+%         while True:
+%           time.sleep(0.5)
+%       except KeyboardInterrupt:
+%         pass
+% \end{pythoncode}
+% \caption{Full example of a \pyRobots application.}
+% \label{mwe}
+% \end{listing}
+%
 
@@
 \paragraph{Stress-test in a nursery}
+\label{croq}
 
@@
     \item path to more complex control tool (T-REX, etc)
+    \item investigate techniques in Python's {\sc asyncio} module
 \end{itemize}

### 2014-09-17 paper.tex
Cleanup fig croquignole expe
Abstract + acknowledgments

@@
 \usepackage{amsmath}
+\usepackage{booktabs}
 
@@
 
-    Blabla
+Presented is \pyRobots, a new software toolkit for the executive control of
+complex robotic systems. Borrowing ideas from previous tools like {\sc
+urbi}~\cite{baillie2005urbi}, it proposes a lightweight Python framework to
+develop (pre-emptive) concurrent and event-based executive controllers.
+
+Designed in a \emph{bottom-up} fashion, out of the need for a practical,
+unobstrusive tool suitable for rapid prototyping of complex interaction
+scenarios, \pyRobots also exposes several simple yet convenient abstractions for
+physical resources management and pose representation.
+
+Finally, and while middleware-agnostic, it features advanced integration with
+the \textit{de-facto} standard ROS middleware.
 
@@
 \begin{tabular}{p{0.9\columnwidth}}
-\hline
+    \toprule
     {\bf Manipulation} \\
      {\tt attachobject}, {\tt basicgive}, {\tt basictake}, {\tt close\_gripper}, {\tt configure\_grippers}, {\tt grab\_gripper}, {\tt handover}, {\tt hide}, {\tt open\_gripper}, {\tt pick}, {\tt put}, {\tt put\_accessible}, {\tt release\_gripper}, {\tt show}, {\tt take} \\
-\hline
+     \midrule
     {\bf Gaze control} \\
      {\tt glance\_to}, {\tt look\_at}, {\tt sweep}, {\tt switch\_active\_stereo\_pair}, {\tt track}, {\tt cancel\_track} \\
-\hline
+     \midrule
     {\bf Navigation} \\
      {\tt carry}, {\tt follow}, {\tt cancel\_follow}, {\tt goto}, {\tt moveclose}, {\tt waypoints} \\
-\hline
+     \midrule
     {\bf Local navigation} \\
      {\tt dock}, {\tt rotate}, {\tt translate} \\
-\hline
+     \midrule
     {\bf Posture configuration} \\
      {\tt extractpose}, {\tt idle}, {\tt manipose}, {\tt movearm}, {\tt rest}, {\tt setpose}, {\tt settorso}, {\tt tuckedpose} \\
-\hline
+     \bottomrule
 \end{tabular}
@@
 
-We brought two Ranger robots to a kindergarten (17 children, mean age: 3.88
-years old, SD=0.65), and after a brief introduction (showing them that removing
-the pacifier would ``wake up'' the robot, and putting it back would make it
-``fall asleep''), we invited the children to freely play with the robots and
-integrate them into their usual games. The experiment lasted 68 minutes.
+We brought two Ranger robots to a kindergarten (17 children, age: M=3.88,
+SD=0.65), and after a brief introduction (showing them that removing the
+pacifier would ``wake up'' the robot, and putting it back would make it ``fall
+asleep''), we invited the children to freely play with the robots and integrate
+them into their usual games. The experiment lasted 68 minutes.
 
@@
 \begin{figure}
-    \resizebox{\columnwidth}{!}{\input{croquignole-expe}}
+    \centering
+    \includegraphics[width=0.9\columnwidth]{croquignole-expe}
     \caption{Events fired and actions started per minute, over the 68 min-long
@@
 
-\section*{Acknowledgment}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Acknowledgments}
 
-This work has been supported by...
+This research was supported by the Swiss National Science Foundation through the
+National Centre of Competence in Research Robotics.
 

### 2014-09-18 paper.tex
Lot of little things

@@
 \newcommand{\eg}{{\textit{e.g.\ }}}
-\newcommand{\pyRobots}{\textsc{pyRobots}\ }
+\newcommand{\pyRobots}{\textsc{pyRobots}}
 
 \title{\LARGE \bf
-    \pyRobots, a toolkit for robot executive control
+    \pyRobots{}, a toolset for robot executive control
 }
@@
 
-Presented is \pyRobots, a new software toolkit for the executive control of
+Presented is \pyRobots{}, a new software toolkit for the executive control of
 complex robotic systems. Borrowing ideas from previous tools like {\sc
@@
 unobstrusive tool suitable for rapid prototyping of complex interaction
-scenarios, \pyRobots also exposes several simple yet convenient abstractions for
+scenarios, \pyRobots{} also exposes several simple yet convenient abstractions for
 physical resources management and pose representation.
@@
 Orchestrating the activity of a robot is a central issue of robotics: \emph{what
-to do when?} Traditionally, in the famous 3-layer architectures~\fixme{ref?},
-this function is implemented in the so-called \emph{deliberative layer} where
-decisions are made based on perceptions and on the current internal state, and
-executed by sending orders to a \emph{functional layer}.
+to do when?} Traditionally, in layered architectures, this function is
+implemented in the so-called \emph{deliberative layer} where decisions are made
+based on perceptions and on the current internal state, and executed by sending
+orders to a \emph{functional layer}.
 
@@
 
-This article introduces our attempt at designing an unobtrusive execution control
-toolkit, that aims at addressing the \emph{practicality} issue: it has been
-designed and implemented from the bottom-up, starting from actual needs when
-running complex robots in largely unpredictable scenarios (typically,
-loosely constrained human-robot interactions), and in the real world.
+This article introduces our attempt at designing an unobtrusive execution
+control toolkit, that aims at addressing the \emph{practicality} issue: it has
+been designed and implemented from the bottom-up, starting from actual needs
+when running large robotic systems (including Willow Garage's PR2 and
+Aldebaran's Nao) in dynamic and partially unpredictable environments (typically,
+loosely constrained human-robot interaction scenarios).
 
 Instead of an \emph{environment} or a \emph{framework}, which would suggest
-strong design and development constraints, \pyRobots can be conceived as a set of
-powerful software helpers to write parallel, event-based, high-level robot
-controllers.
+rigid design and development guidelines, \pyRobots{} can be conceived as a set
+of software helpers to write parallel, event-based, high-level robot
+controllers.  It is written for and in Python, which ensure familiarity, fast
+development cycles and broad support for interfacing with existing robots and
+middlewares. \pyRobots{} is indeed \emph{not} a middleware: it purposefully
+provides no mechanisms to connect to other modules or components. It instead
+\emph{relies} on one (or several) middleware to actually communicate with the
+robot and perform actions.
 
-It is written for and in Python, which ensure familiarity, fast development
-cycles and broad support for interfacing with existing robots and middlewares.
-This must be emphasized: \pyRobots is \emph{not} a middleware. It purposefully
-provides no mechanisms to connect to other modules or components. It instead
-\emph{uses} one (or several) middleware to actually communicate with the robot
-and perform actions.
-
-One the other hand, \pyRobots does not provide formal models or verifiability:
-in its current state, it does not attempt to contribute in this field. It
-focuses instead on providing unobtrusive tools to control robots in environments
-that require carrying out many tasks in parallel, in an unpredictable,
-event-based manner.
+Also, \pyRobots{} does not provide formal models or verifiability: in its
+current state, it does not attempt to contribute in this field. It focuses
+instead on providing a lightweight tool to control robots in environments that
+require carrying out many tasks in parallel, in an unpredictable, event-based
+manner. As such, \pyRobots{} is focusing on use-cases ranging from prototyping
+to mid-size experiments.
 
@@
 supervise the robot behaviour, and, expectedly, many projects have attempted to
-design and provide tools for this task.  We focus this review on \emph{tools
-dedicated to the explicit implementation of high-level behaviours}, excluding
-complete control stacks (that would typically include a middleware and hardware
-abstractions), meta-tools (like component generators or model checkers) and
-(semi-)automatic techniques (like learning by demonstration).  In contrast with
-middlewares for which the robotic community relies on a few \emph{de-facto}
-standards (ROS, YARP, OpenRTM), we want to underline that no such widely-used
-softwares exist for execution control, and we believe that most of today's robot
-behaviours are implemented as \emph{ad-hoc} scripts.
+design and provide tools for this task. Degroote provides an in-depth review of
+the existing techniques and tools in~\cite{degroote2012architecture}. We focus
+this review on tools dedicated to the explicit implementation of high-level
+behaviours, excluding complete control stacks (that would typically include a
+middleware and hardware abstractions), meta-tools (like component generators or
+model checkers) and (semi-)automatic techniques (like emerging behaviours or
+learning by demonstration).  In contrast with middlewares for which the robotic
+community relies on a few \emph{de-facto} standards (ROS, YARP, OpenRTM), we
+want to underline that no such widely-used softwares exist for execution
+control, and we believe that most of today's robot behaviours are implemented as
+\emph{ad-hoc} scripts.
 
@@
 RPL~\cite{mcdermott1993reactive}, Golog~\cite{levesque1997golog} (and its
-derivative like {\sc ReadyLOG}~\cite{ferrein2007robot} or, more recently,
-URBI~\cite{baillie2005urbi}; or what Joyeux calls \emph{embedded}
+derivatives like {\sc ReadyLOG}~\cite{ferrein2007robot}) or, more recently,
+{\sc urbi}~\cite{baillie2005urbi}; or what Joyeux calls \emph{embedded}
 DSL~\cite{joyeux2011robot}: extensions of existing languages (typically, dynamic
@@
 roby}~\cite{joyeux2009plan} (based on Ruby), {\sc
-teer}~\footnote{\url{http://wiki.ros.org/executive_teer}} (based on Python), are
+teer}~\cite{magnenat2012teer} (based on Python), are
 examples of such eDSL.
@@
 
-\pyRobots positions itself in this landscape as a lightweight eDSL (in a similar
+\pyRobots{} positions itself in this landscape as a lightweight eDSL (in a similar
 fashion to {\sc roby} or {\sc teer}), that focuses on high-level behaviours
@@
 the robot software stack, using Python as ubiquitous binding language. Similar
-to URBI, it focuses on transparent preemptive concurrency and event-based programming.
-Also, \pyRobots' model favours hierarchical task composition over plans, and does not
+to {\sc urbi}, it focuses on transparent preemptive concurrency and event-based programming.
+Also, \pyRobots{}' model favours hierarchical task composition over plans, and does not
 provide by itself any task planning service, delegating this to external tools.
@@
 
-Before introducing and explaining step-by-step a complete working example in the
-next section, it is useful to first introduce the main concepts of \pyRobots,
-and provide an overview of the features and mechanisms. Since \pyRobots is an
-open-source software, the interested reader can also refer to the source code,
-available here:~\url{https://github.com/chili-epfl/pyrobots}.
-
-\begin{itemize}
-    \item Lock specific resources with a simple \python|@lock(...)| in front of the
-        actions. When starting, actions will wait for resources to be
-        available if needed.
-    \item Supports compound resources (like \python|WHEELS == LEFTWHEEL + RIGHTWHEEL|)
-    \item Poses are managed explicitly and can easily be
-        transformed from one reference frame to another one
-        (integrates with ROS TF when available).
-    \item Extensive logging support to debug and replay
-        experiments.
-\end{itemize}
+This section introduces the main concepts of \pyRobots{},
+and provide an overview of the features and mechanisms. Since \pyRobots{} is an
+open-source software, the interested reader can also refer to the source code
+and documentation, available from~\url{https://github.com/chili-epfl/pyrobots}.
 
@@
 
-\pyRobots core concepts are simple: on one hand, the user creates a
+\pyRobots{} core concepts are simple: on one hand, the user creates a
 \textbf{robot} as an instance of a \python|GenericRobot|, that encapsulates the
@@
 (typically a set proxies to the sensors, or a connector to a database, etc.,
-accessed through a dictionary-like object).
+accessed through a key-value container -- a dictionary-like object).
 
 On the other hand, the user creates as many high-level \textbf{actions} as
-desired, as regular Python functions. Actions are automatically added to the
+desired, as regular Python functions, only annotated with the \python|@action|
+decorator. Actions are automatically added to the
 \textbf{robot}, and can access its state and controllers to perform actual
@@
 Finally, the user can declare arbitrary \textbf{resources}, usually
-corresponding to the robot components (for instance, the wheels, the camera,...)
-and guarantee exclusive access to the required resource to specific actions.
+corresponding to the robot physical components (for instance, the wheels, the
+camera) and guarantee exclusive access to the required resource to specific
+actions.
 
@@
 
-\emph{Actions} in \pyRobots are always asynchronous. They are implemented as
+\emph{Actions} in \pyRobots{} are asynchronous by default. They are implemented as
 \emph{futures} (also known as \emph{promises}) and therefore executed in
 independent threads. To the developer, however, they appear as regular Python
-function, simply annotated with the \python|@action| decorator, and are invoked
+function annotated with the \python|@action| decorator, and are invoked
 like any other function.  Since they are asynchronous, these functions may
-perfectly block or even never return (typically useful to implement lasting
+perfectly block or even never return (typically useful to implement continuous
 background behaviours).
@@
 
-\paragraph{Signaling} Behaviours often need to be paused of cancelled.  This
-raises an issue that needs to be addressed in robotics: before suspending an
-action, one often wants to first bring back the robot into some form of rest
+\emph{Signaling:} Behaviours often need to be paused or cancelled.  This raises
+an issue that specifically needs to be addressed in robotics: before suspending
+an action, one often wants to first bring back the robot into some form of rest
 state (\eg one would not want to suddenly cancel a \python|walk| action in the
-middle of a step without first bringing back the foot to the ground).
-
-\pyRobots addresses this issue through signals, and extends standard
-Python thread into \emph{signaling threads}, relying on Python's usual syntax
-for exception handling. Listing~\ref{lst:signals}
-illustrates this mechanism. The action is instantiated and started at line 8,
-and cancelled at line 10. This send a \python|ActionCancelled| signal to the
-action that the user can handle using an \python|except| statement.
+middle of a step without first bringing back the foot to the ground). \pyRobots{}
+addresses this issue through signals, and extends standard Python thread into
+\emph{signaling threads}, relying on Python's usual syntax for exception
+handling. Listing~\ref{lst:signals} illustrates this mechanism. The action is
+instantiated and started at line 8, and cancelled at line 10. This send a
+\python|ActionCancelled| signal to the action that the user can handle using an
+\python|except| statement.
 
@@
     @action
-    def my_action(robot):
+    def safe_walk(robot):
       try:
-        robot.do_something_long()
+        robot.walk()
       except ActionCancelled:
@@
 
-    action = robot.my_action()
+    action = robot.safe_walk()
     time.sleep(1)
@@
 \end{pythoncode}
-\caption{Handling a cancellation signal}
+\caption{\textbf{Handling a cancellation signal} After one second, the
+\python|safe_walk| action is cancelled. This sends the signal
+\python|ActionCancelled| to the action, that can appropriately terminate.}
 \label{lst:signals}
@@
 given ways to code transition behaviours when pausing and resuming) and is not
-currently supported by \pyRobots.
+currently supported by \pyRobots{}.
 
@@
 
-Sharing the same syntax with URBI, events in \pyRobots can by monitored through
-the methods \python|on| (one shot event) and \python|whenever| (continuous event
-monitoring). Event conditions can be either a predicate (that must take as
-single argument the robot instance) or a pair of $\{\text{state member},
-\text{condition}\}$ with condition in $\{\text{value}=x, \text{become}=x,
-\text{below}=x, \text{above}=x, \text{decrease}=x, \text{increase}=x\}$. An
-example of event usage can be seen at line 67 of listing~\ref{mwe}.
+Sharing the same syntax with {\sc urbi}, events in \pyRobots{} can by monitored
+through the methods \python|on| (one shot event) and \python|whenever|
+(continuous event monitoring). Event conditions can be either a predicate (that
+must take as single argument the robot instance) or a pair of $\{\text{state
+key}, \text{condition}\}$ with condition in $\{\text{value}=x, \text{become}=x,
+\text{below}=x, \text{above}=x, \text{decrease}=x, \text{increase}=x\}$.
+Examples of event usage can be seen at lines 33--38 of
+listing~\ref{croquignole_no_move}.
 
-When the event condition is verified, the callback (usually, an action) is
-invoked from the event monitor thread.
+Each event monitor lives in its own separate thread, and when the event is fired,
+the callback (usually, an action) is invoked from the event monitor thread
+itself.
 
@@
 
-Usually, physical robot resources (mainly actuators, but possibly sensors as
-well) require exclusive use (it is seldom desirable that two different actions
-control the wheels of a robot at the same time).
+Physical robot resources (mainly actuators, but possibly sensors as
+well) usually mandate exclusive use (it is seldom desirable that two different actions
+control the wheels of a robot at the same time). \pyRobots{} provides a resource
+locking mechanism that relies on the \python|@lock| annotation: an action
+annotated with \python|@lock(ARMS)| (Listing~\ref{lst:resources}, l.6) will
+acquire (or wait for) the \python|ARM| resource when being executed.
 
-To leave full control regarding the required resource modeling granularity, the
-programmer define them himself, either as stand-alone resources or as compound
-resources (listing~\ref{lst:resources}).
+This mechanism is generic and \pyRobots{} makes no assumptions concerning the
+actual resources available/meaningful to a specific robot. The
+programmer define them himself, either as stand-alone or as compound
+resources (listing~\ref{lst:resources}, l.1--3).
 
@@
     robot.wave_hand() # waits until lift_box is over
-    robot.scratch_head() # skipped if lift_box 
-                         # is still running
+    robot.scratch_head() # skipped if lift_box or
+                         # wave_hand are still running
 
 \end{pythoncode}
-\caption{Resource usage is defined at the action-level, through annotations.}
+\caption{\textbf{Resource locking} Resource usage is defined at the
+action-level, through annotations.}
+
 \label{lst:resources}
@@
 
-The user then annotates the actions with the (possibly multiple) resources they
-require (lines 6, 11 and 16 of listing~\ref{lst:resources}). When executed, the
+A given action can lock several distinct resources (with the corresponding
+number of \python|@lock| annotations. When executed, the
 action acquires a lock on the resource (and all dependent resources in the case
-of a compound resource) which is released upon action completion. In the
-meantime, other actions requiring the same resource will either wait until the
-resource is release (line 21) or be skipped (line 22) if the \python|@lock|
-annotation specifies not to wait (line 16).
+of a compound resource) which is released upon action completion. If the lock is
+already acquired, two strategies are possible: either wait or \emph{fail fast}
+(\python|@lock(..., wait=False)|, l.16).
 
@@
 
-\pyRobots offers a similar approach \emph{within} the control script, adding a
+\pyRobots{} offers a similar approach \emph{within} the control script, adding a
 pinch of syntactic sugar to make poses and frame management transparent for the
@@
 qx, qy, qz, qw, frame\}$ or by a dictionary that defines any of these 8 values.
-Internally, \pyRobots normalizes and stores the poses in their full 8-tuple
+Internally, \pyRobots{} normalizes and stores the poses in their full 8-tuple
 expansion.
@@
 \end{pythoncode*}
-\caption{Examples of pose normalization.}
+\caption{Examples of \textbf{pose normalization}. Poses can be transformed to
+other reference frames with the \python|inframe| method (which implicitly performs
+normalization if needed).}
 \label{lst:poses}
@@
 
-These transformations are directly computed, but \pyRobots also takes advantage
+These transformations are directly computed, but \pyRobots{} also takes advantage
 of middleware-specific infrastructure (when available) to provide known frames.
 ROS TF, Aldebaran's {\tt NAOqi} frame conventions, etc. are used when available
-through specializations of the \pyRobots \python|PoseManager| class.
+through specializations of the \pyRobots{} \python|PoseManager| class.
 
@@
 Running robots in real-world, \ie dynamic and unpredictable, environments
-entails repetitive coding-testing-debugging development cycles. \pyRobots
+entails repetitive coding-testing-debugging development cycles. \pyRobots{}
 provides several tools to support this development phase. A special attention
@@
 
-\pyRobots logs are designed to allow full \emph{post-mortem} replay, and we
+\pyRobots{} logs are designed to allow full \emph{post-mortem} replay, and we
 provide a graphical interface to display the actions and resource usage over
@@
         \includegraphics[width=0.9\columnwidth]{log}
-        \caption{Screenshot of \pyRobots log analyser, processing the log of the
+        \caption{Screenshot of \pyRobots{} log analyser, processing the log of the
         experiment presented at section~\ref{croq}}
@@
 % \end{pythoncode}
-% \caption{Full example of a \pyRobots application.}
+% \caption{Full example of a \pyRobots{} application.}
 % \label{mwe}
@@
 
-The development of \pyRobots started in 2011, driven by the requirements of a
+The development of \pyRobots{} started in 2011, driven by the requirements of a
 18-min long theatre performance between a human actor and a robot. Since that time, the toolkit has
@@
 \end{center}
-\caption{Examples of \pyRobots actions, mostly extracted from experiments involving
+\caption{Examples of \pyRobots{} actions, mostly extracted from experiments involving
     the PR2 robot.}
@@
 
-\paragraph{Roboscopie} The first complex deployment scenario for \pyRobots was
+\subsubsection{Roboscopie} The first complex deployment scenario for \pyRobots{} was
 indeed a theatre performance entitled
@@
 the need to control a complex robot (a Willow Garage PR2) in a structured way,
-as required for a live performance with a human\footnote{The original \pyRobots
+as required for a live performance with a human\footnote{The original \pyRobots{}
 script used for the performance can be accessed
@@
 
-A total of 47 \pyRobots actions were written for this event
+A total of 47 \pyRobots{} actions were written for this event
 (Table~\ref{pyrobots_actions} gives a few examples), and the toolkit proved to
@@
 
-\pyRobots also allowed us to retain the \emph{rapid prototyping} convenience of a
+\pyRobots{} also allowed us to retain the \emph{rapid prototyping} convenience of a
 script language, essential for the robot to ``practise'' and adapt to the
@@
 
-\paragraph{Stress-test in a nursery}
+\subsubsection{Stress-test in a nursery}
 \label{croq}
@@
 
-Software-wise, all the user-facing behaviours are implemented with \pyRobots,
+Software-wise, all the user-facing behaviours are implemented with \pyRobots{},
 and the communication with the hardware is abstracted through a mix of ROS
@@
         \caption{Infants playing freely (!) with an EPFL's \emph{Ranger} robot.
-        The behaviour of the robot is controlled with \pyRobots.}
+        The behaviour of the robot is controlled with \pyRobots{}.}
         \label{expe-nursery}
@@
 
-        robot.every("pacifier", becomes = True).do(on_pacifier)
-        robot.every("pacifier", becomes = False)
+        robot.whenever("pacifier", becomes = True)
+                                    .do(on_pacifier)
+        robot.whenever("pacifier", becomes = False)
                                     .do(on_pacifier_removed)
-        robot.every("scale", increase = 0.3).do(on_toy)
-        robot.every("bumper", becomes = True).do(on_bumper)
+        robot.whenever("scale", increase = 0.3).do(on_toy)
+        robot.whenever("bumper", becomes = True).do(on_bumper)
 

### 2014-09-18 paper.tex
Rewritten section on pose management

@@
 task, mainly because it seems that every single library, middleware or robotic
-platform rely on its own conventions (rotation representations, in particular,
-are a constant source of frustration, having to jump all the time between
-quaternions, one of the 23 Euler conventions, rotation matrices, Rodrigues,
-yaw-pitch-roll notation...). Within the ROS ecosystem, the Transform Library
-(tf,~\cite{foote2013tf}) has played a major role to make it convenient to work
+platform relies on its own conventions (rotation representations, in particular,
+are a common source of inconsistencies -- quaternions, (the many) Euler
+conventions, rotation matrices, Rodrigues, yaw-pitch-roll notations being all
+rather frequent). Within the ROS ecosystem, the Transform Library ({\sc
+tf},~\cite{foote2013tf}) has played a major role to make it convenient to work
 with frames, and poses can always be expressed and shared in the most convenient
-local frame, tf taking care of required intermediate transformation.
+local frame, {\sc tf} taking care of required intermediate transformation.
 
-\pyRobots{} offers a similar approach \emph{within} the control script, adding a
-pinch of syntactic sugar to make poses and frame management transparent for the
-high-level action controller.
-
-As illustrated in listing~\ref{lst:poses}, poses can be flexibly defined by a
-simple frame name, by a triple $\{x, y, z\}$ (default frame {\tt map} is then
-assumed), by a 4-tuple $\{x, y, z, frame\}$, by a 6-tuple $\{x, y, z, rx, ry,
-rz\}$ (using $sxyz$ Euler convention), by a 7-tuple $\{x, y, z, rx, ry, rz,
-frame\}$, by a 7-tuple $\{x, y, z, qx, qy, qz, qw\}$, by a 8-tuple $\{x, y, z,
-qx, qy, qz, qw, frame\}$ or by a dictionary that defines any of these 8 values.
-Internally, \pyRobots{} normalizes and stores the poses in their full 8-tuple
-expansion.
+\pyRobots{} pose management builds on this idea to offer flexible pose
+representations \emph{within} the control scripts. Everytime a 6D pose needs to
+be specified (like a navigation goal or a target pose for an arm), it can be
+expressed as a simple frame name, by a triple $\{x, y, z\}$ (default frame {\tt
+map} is then assumed), by a 4-tuple $\{x, y, z, frame\}$, by a 6-tuple $\{x, y,
+z, rx, ry, rz\}$ (using $sxyz$ Euler convention), by a 7-tuple $\{x, y, z, rx,
+ry, rz, frame\}$, by a 7-tuple $\{x, y, z, qx, qy, qz, qw\}$, by a 8-tuple $\{x,
+y, z, qx, qy, qz, qw, frame\}$ or by a dictionary that defines any subset of
+these 8 values.  Internally, \pyRobots{} normalizes and stores the poses in
+their full 8-tuple expansion, as illustrated in listing~\ref{lst:poses}.
 
@@
 
-These transformations are directly computed, but \pyRobots{} also takes advantage
-of middleware-specific infrastructure (when available) to provide known frames.
-ROS TF, Aldebaran's {\tt NAOqi} frame conventions, etc. are used when available
-through specializations of the \pyRobots{} \python|PoseManager| class.
+Required transformations are computed by \pyRobots{}, and the programmer can
+provide new ``sources'' of frame transformations through the instanciation of
+\python|FrameProviders| that takes advantage of robot- and middleware-specific
+infrastructures (for example, ROS {\sc tf}, Aldebaran's {\tt NAOqi} frame
+conventions, etc.).
 

### 2014-09-18 paper.tex
Completed studies section

@@
 desired, as regular Python functions, only annotated with the \python|@action|
-decorator. Actions are automatically added to the
-\textbf{robot}, and can access its state and controllers to perform actual
-physical actions.
+decorator. Actions are automatically added to the \textbf{robot}, and can access
+its state and controllers to perform actual physical actions.
 
@@
 
-Actions can also call each other, allowing natural decomposition of complex
-tasks into sub-actions and atomic actions.
+Actions can also call each other (nested actions), allowing natural
+decomposition of complex tasks into sub-actions and atomic actions (we note
+hereafter $depth$ the nesting level of an action, \ie the maximum number of
+calls to sub-actions until an atomic action is reached).
 
@@
 
+A given action can lock several distinct resources (with the corresponding
+number of \python|@lock| annotations. When executed, the
+action acquires a lock on the resource (and all dependent resources in the case
+of a compound resource) which is released upon action completion. If the lock is
+already acquired, two strategies are possible: either wait or \emph{fail fast}
+(\python|@lock(..., wait=False)|, l.16).
+
 \begin{listing}[H]
@@
 \end{listing}
-
-A given action can lock several distinct resources (with the corresponding
-number of \python|@lock| annotations. When executed, the
-action acquires a lock on the resource (and all dependent resources in the case
-of a compound resource) which is released upon action completion. If the lock is
-already acquired, two strategies are possible: either wait or \emph{fail fast}
-(\python|@lock(..., wait=False)|, l.16).
 
@@
 The development of \pyRobots{} started in 2011, driven by the requirements of a
-18-min long theatre performance between a human actor and a robot. Since that time, the toolkit has
-iteratively matured, and has been deployed and used on several projects and
-platforms. We first briefly recap these deployments with references to the
-corresponding publications, and then present our latest study, in a nursery,
-that recently acted as a stress-test for the library.
+18min long theatre performance between a human actor and a robot. Since that
+time, the toolkit has iteratively matured, and has been deployed and used on
+several projects and platforms. We first briefly recap these deployments with
+references to the corresponding publications, and then present our latest study,
+in a kindergarten, that recently acted as a stress-test for the library.
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{lll}
+        \toprule
+        Study                 & Robot & Middleware \\ \midrule
+        Roboscopie            & PR2       & ROS, {\sc GenoM} \\
+        Interactive Grounding & PR2, LAAS-CNRS's Jido & ROS, {\sc GenoM} \\
+        CoWriter              & Nao       & ROS, NAOqi  \\
+        Ranger at kindergarten & EPFL's Ranger & ROS, aseba  \\
+        \bottomrule
+    \end{tabular}
+    \caption{Overview of significant experiments and studies where \pyRobots{}
+    has been used.}
+\end{table}
+
+\subsubsection{Roboscopie} The first complex deployment scenario for \pyRobots{} was
+a theatre performance entitled \emph{Roboscopie}~\cite{lemaignan2012roboscopie}.
+The toolkit was created out of the need to control a complex robot (a Willow
+Garage PR2) in a structured way, as required for a live performance with a
+human\footnote{The original \pyRobots{} script used for the performance can be
+accessed from~\url{http://www.laas.fr/roboscopie}. Note that this script does
+not reflect anymore the toolkit usage.}
+
+A total of 47 \pyRobots{} actions were written for this event
+(Table~\ref{pyrobots_actions} gives a few significant examples), and the toolkit
+proved to be especially convenient to interleave calls to both ROS-based and
+{\sc GenoM}-based~\cite{mallet2010genom3} components through their respective
+Python bindings.
+
+\pyRobots{} also allowed us to retain the \emph{rapid prototyping} convenience of a
+script language, essential for the robot to ``practise'' and adapt to the
+vision of the director.
 
@@
 \end{center}
-\caption{Examples of \pyRobots{} actions, mostly extracted from experiments involving
-    the PR2 robot.}
+\caption{Examples of \pyRobots{} actions (extracted from experiments involving
+the PR2 robot).}
 
@@
 
-\subsubsection{Roboscopie} The first complex deployment scenario for \pyRobots{} was
-indeed a theatre performance entitled
-\emph{Roboscopie}~\cite{lemaignan2012roboscopie}. The toolkit was created out of
-the need to control a complex robot (a Willow Garage PR2) in a structured way,
-as required for a live performance with a human\footnote{The original \pyRobots{}
-script used for the performance can be accessed
-from~\url{http://www.laas.fr/roboscopie}. Note that this script does not
-reflect anymore the toolkit usage.}
+\subsubsection{Interactive Grounding} This study on perspective-taking and grounding
+in human-robot interaction~\cite{lemaignan2013talking} is one of the several
+experiments that was built on top of the LAAS-CNRS cognitive
+architecture~\cite{lemaignan2014human}. In this large architecture, \pyRobots{}
+is one of the two available executive controllers, and integrates with a task
+planner, a symbolic knowledge base, several motion planner and a natural
+language processor. \pyRobots{} was favoured over the second execution
+controller (based on PRS) for smaller studies, as a more powerful alternative to
+bare \textit{ad-hoc} scripts.
 
-A total of 47 \pyRobots{} actions were written for this event
-(Table~\ref{pyrobots_actions} gives a few examples), and the toolkit proved to
-be especially convenient to interleave calls to both ROS-based and {\sc
-GenoM}-based\footnote{{\sc GenoM} is software component as well as a middleware
-developed at LAAS-CNRS} components through their respective Python bindings.
-
-\pyRobots{} also allowed us to retain the \emph{rapid prototyping} convenience of a
-script language, essential for the robot to ``practise'' and adapt to the
-vision of the director.
-
-\begin{itemize}
-    \item PR2, with ROS and pocolibs
-    \item Nao, with ROS and NAOqi
-    \item Ranger with Aseba and ROS
-\end{itemize}
-
-\begin{itemize}
-    \item roboscopie
-    \item LAAS architecture
-    \item CoWriter
-    \item Ranger
-\end{itemize}
+\subsubsection{CoWriter} The {\sc CoWriter} project~\cite{hood2015those}
+(Figure~\ref{expe-cowriter}) involves an Aldebaran Nao robot and young students
+in a \emph{learning by teaching} interaction: the children have to teach the
+robot how to handwrite. This experiment involved a dual ROS/NAOqi environment,
+bridged within \pyRobots{}. The high-level behaviour was implemented as a finite
+state machine ({\sc SMACH}) whose state corresponded to \pyRobots{} actions.
+While the execution controller was subsequently simplified (towards a tighter
+interaction scenario), this illustrates how \pyRobots{} -- relying on the
+ubiquitous Python language -- can possibly blend into and interact with existing
+tools.
 
@@
 
-\subsubsection{Stress-test in a nursery}
+\subsubsection{Stress-test in a kindergarten}
 \label{croq}
@@
 removable pacifier (used as a soft on/off switch by the children) and it is
-covered with capacitive tactile panels. The wooden panels cover 186 RGB LEDs,
+covered with capacitive tactile panels. The wooden panels hide 186 RGB LEDs,
 enabling various light patterns, it can play sounds, the two eyes can be fully
@@
 such as they switch to freewheel as soon as the robot is pushed by the children.
-The robot is powered by an embedded Gumstix Overo (ARM7l) board running a custom
+The robot is powered by an embedded Gumstix Overo board (ARM7l) running a custom
 Linux image, assisted by three custom microcontroller boards (for low-level
@@
         \caption{Infants playing freely (!) with an EPFL's \emph{Ranger} robot.
-        The behaviour of the robot is controlled with \pyRobots{}.}
+        The behaviours of the robot are implemented via \pyRobots{}.}
         \label{expe-nursery}
@@
 
-We brought two Ranger robots to a kindergarten (17 children, age: M=3.88,
+We brought two such Ranger robots to a kindergarten (17 children, age: M=3.88,
 SD=0.65), and after a brief introduction (showing them that removing the
@@
 (Listing~\ref{croquignole_no_move}): two background actions
-(\python|background_blink| and \python|look_at_touches|) run continuously, and 4
-events are monitored (pacifier added/removed, toy added and bumper hit).
+(\python|background_blink| and \python|look_at_touches|) run continuously, and 5
+events are monitored (pacifier added/removed, toy added/removed and bumper hit).
 
@@
     def on_pacifier(robot):
-        robot.look_at_lolette()
+        robot.look_at_pacifier()
         robot.blink()
@@
 \end{pythoncode}
-\caption{Source of the high-level behaviours running on the robots during the
-nursery pilot (some behaviours like battery management have been omitted for
+\caption{Simplified source of the high-level behaviours running on the robots during the
+kindergarten study (some behaviours like battery management have been omitted for
 clarity).}
@@
 
+Figure~\ref{croq_actions} shows the average number of events fired and action
+(including sub-actions) \emph{started} per minute, for the whole experiment. The
+diagram reveals several peaks above 180 actions/minutes with a global average of
+70 actions/minutes, which was surprisingly high considering the simple
+behavioural scheme we proposed and the limited actions' depth
+($depth_{max}=depth_{\tt look\_at\_touches}=3$) . As a safety measure, we also
+set for this study a hard limit to the maximum amount of actions allowed to run
+in parallel (20 actions): again, to our surprise, this limit was hit 31 times
+during the experiment.
+
 \begin{figure}
@@
     \includegraphics[width=0.9\columnwidth]{croquignole-expe}
-    \caption{Events fired and actions started per minute, over the 68 min-long
+    \caption{Events fired and actions started per minute, over the 68min long
     experiment. The robot sustains an average of 70 actions triggered by minute over
     more than an hour, with peaks at 200 actions/minute.}
+    \label{croq_actions}
+\end{figure}
 
-\end{figure}
+The results of this stress-test provide a reasonable baseline for the
+performances in terms of parallelism and debug/introspection capabilities that
+can be expected from an executive controller.
 

### 2014-09-18 paper.tex
Conclusion

@@
 
-\subsection{Summary and main contributions}
+We have presented here \pyRobots{}, a novel Python library supporting
+development of high-level executive controllers for robotics.
 
-\begin{itemize}
-    \item strives to simplicity (from a user perspective)
-    \item Middleware independent, real-world deployments with with ROS, GenoM, NAOqi, aseba
-\end{itemize}
+Its main features include a lightweight, unobtrusive design based on code
+annotations; a pre-emptive concurrency model allowing for proper, explicit
+handling of task cancellation; support for event-based programming; a flexible
+robot resources locking mechanism and a built-in support for 6D pose
+manipulation and transformation.
 
-\subsection{Future work}
+\pyRobots{} does not aim at replacing large execution control frameworks. It
+proposes instead an accessible set of tools to improve upon the all-too-common
+\emph{ad-hoc} scripts, while remaining middleware agnostic.
 
-\begin{itemize}
-    \item control mixing policies?
-    \item control over action priorities
-    \item path to more complex control tool (T-REX, etc)
-    \item investigate techniques in Python's {\sc asyncio} module
-\end{itemize}
+The article also reports on several experimental deployments on four different
+kind of robots using a mixture of four different middlewares.
+
+While already four years old, several possible developments would offer
+interesting perspectives for this tool, including (a)~the support to suspend (and
+then resume) actions (which requires defining a satisfying syntax to declare
+suspend/resume behaviours within actions); (b)~mechanisms to define action
+priorities; (c)~investigate recent work on Python's {\tt asynchio} mechanisms
+(that provide reference implementation of co-routines in Python).
+
+Being available as an open-source platform, documented and easy to install, we
+believe that these development objectives may materialize soon, as well-defined
+use-cases emerge from the library's users.
 

### 2014-09-18 paper.tex
Many little things

@@
 
-Domain-specific languages (DSL), in the widest sense, represent the largest body
-of literature on high-level behaviour programming. They can either be ``actual''
-domain-specific languages like PRS~\cite{ingrand1996prs},
+Domain-specific languages (DSL), in the broadest sense, represent the largest
+body of literature on high-level behaviour programming. They can either be
+actual domain-specific languages like PRS~\cite{ingrand1996prs},
 RPL~\cite{mcdermott1993reactive}, Golog~\cite{levesque1997golog} (and its
-derivatives like {\sc ReadyLOG}~\cite{ferrein2007robot}) or, more recently,
-{\sc urbi}~\cite{baillie2005urbi}; or what Joyeux calls \emph{embedded}
+derivatives like {\sc ReadyLOG}~\cite{ferrein2007robot}) or, more recently, {\sc
+urbi}~\cite{baillie2005urbi}; or what Joyeux calls \emph{embedded}
 DSL~\cite{joyeux2011robot}: extensions of existing languages (typically, dynamic
@@
 robots. {\sc cram}~\cite{beetz2010cram} (based on Lisp and Prolog), {\sc
-roby}~\cite{joyeux2009plan} (based on Ruby), {\sc
-teer}~\cite{magnenat2012teer} (based on Python), are
-examples of such eDSL.
+roby}~\cite{joyeux2009plan} (based on Ruby), {\sc teer}~\cite{magnenat2012teer}
+(based on Python), are examples of such eDSL.
 
@@
 fashion to {\sc roby} or {\sc teer}), that focuses on high-level behaviours
-implementation. It seeks unobstrusive interaction with the other components of
+implementation. It seeks unobtrusive interaction with the other components of
 the robot software stack, using Python as ubiquitous binding language. Similar
@@
 the robot) and uniform manipulation of 6D poses (that integrates transparently
-with other tools like ROS's TF when available).
+with other tools like ROS's {\sc tf} when available).
 
@@
 
-\begin{listing}[H]
+\begin{listing}
 \begin{pythoncode}
@@
 \python|safe_walk| action is cancelled. This sends the signal
-\python|ActionCancelled| to the action, that can appropriately terminate.}
+\python|ActionCancelled| to the action, that can appropriately handle it inside
+the \python|except| block.}
 \label{lst:signals}
@@
 
-\begin{listing}[H]
+\begin{listing}
     \begin{pythoncode*}{linenos=false, xleftmargin=0em}
@@
 
-Using helper methods like \python|robot.poses.inframe(<pose>, <other frame>)| or
-\python|robot.poses.pantilt(<pose>, <ref frame>)|, the programmer can then
+Using helper methods like \python|inframe(<pose>, <other frame>)| or
+\python|pantilt(<pose>, <ref frame>)|, the programmer can then
 transform a given pose into a different frame or different representation.
@@
         \centering
-        \includegraphics[width=0.9\columnwidth]{log}
+        \includegraphics[width=0.8\columnwidth]{log}
         \caption{Screenshot of \pyRobots{} log analyser, processing the log of the
@@
 While the execution controller was subsequently simplified (towards a tighter
-interaction scenario), this illustrates how \pyRobots{} -- relying on the
-ubiquitous Python language -- can possibly blend into and interact with existing
+interaction scenario), this illustrates how \pyRobots{}, relying on the
+ubiquitous Python language, can possibly blend into and interact with existing
 tools.

### 2014-09-18 paper.tex
Lot of little things, including making sure it fits into 6 pages

@@
 
-temporal planning/temporal constraints: T-REX~\cite{mcgann2007trex} -> sensing-planning-acting
-
-ROAR: ~\cite{degroote2011roar}
+The T-REX architecture~\cite{mcgann2007trex} partitions the execution control
+problem into several agents, each one composed of a (temporal) planner and an
+execution layer. At run-time, the different agents are constantly synchronised
+to maintain the consistency of the global plan. ROAR~\cite{degroote2011roar}
+augments this idea with a logic mechanism allowing the system to automatically
+select and start agents depending on the task at hand.
 
@@
 
-Projects like PRS, Golog or {\sc cram} have large scopes:
+\pyRobots{} positions itself in this landscape as a lightweight eDSL (in a
+similar fashion to {\sc roby} or {\sc teer}), that focuses on high-level
+behaviours implementation. It seeks unobtrusive interaction with the other
+components of the robot software stack, using Python as ubiquitous binding
+language. Similar to {\sc urbi}, it focuses on transparent pre-emptive
+concurrency and event-based programming.  Also, \pyRobots{}' model favours
+hierarchical task composition over plans, and does not provide by itself any
+task planning service, delegating this to optional external tools.
 
-\pyRobots{} positions itself in this landscape as a lightweight eDSL (in a similar
-fashion to {\sc roby} or {\sc teer}), that focuses on high-level behaviours
-implementation. It seeks unobtrusive interaction with the other components of
-the robot software stack, using Python as ubiquitous binding language. Similar
-to {\sc urbi}, it focuses on transparent preemptive concurrency and event-based programming.
-Also, \pyRobots{}' model favours hierarchical task composition over plans, and does not
-provide by itself any task planning service, delegating this to external tools.
-
-\subsection{Guiding principles}
-
-\begin{itemize}
-    \item Use a de-facto standard language (Python) to ease adoption (familiar to
-        many researchers, large software ecosystem, excellent integration with
-        most robotic softwares, in particular common middlewares like ROS or
-        YARP)
-    \item should be as lightweight/transparent as possible (in particular from a
-        syntax perspective) so that the
-        programmer can focus on the programming of the behaviours (I'm
-        watching you, SMACH)
-    \item should enforce as few design choices as possible. 'convention over
-        configuration' as often as possible
-\end{itemize}
-
-\subsection{Main features}
-
-Parallelism (all \emph{actions} are asynchronous) and event-based programming
-are at the core.
-
-It also offers support for robot resources management (locking of ``parts'' of
-the robot) and uniform manipulation of 6D poses (that integrates transparently
-with other tools like ROS's {\sc tf} when available).
-
+%\subsection{Guiding principles}
+%
+%\begin{itemize}
+%    \item Use a de-facto standard language (Python) to ease adoption (familiar to
+%        many researchers, large software ecosystem, excellent integration with
+%        most robotic softwares, in particular common middlewares like ROS or
+%        YARP)
+%    \item should be as lightweight/transparent as possible (in particular from a
+%        syntax perspective) so that the
+%        programmer can focus on the programming of the behaviours (I'm
+%        watching you, SMACH)
+%    \item should enforce as few design choices as possible. 'convention over
+%        configuration' as often as possible
+%\end{itemize}
+%
+%\subsection{Main features}
+%
+%Parallelism (all \emph{actions} are asynchronous) and event-based programming
+%are at the core.
+%
+%It also offers support for robot resources management (locking of ``parts'' of
+%the robot) and uniform manipulation of 6D poses (that integrates transparently
+%with other tools like ROS's {\sc tf} when available).
+%
 \section{Concepts and Overview}
@@
         \centering
-        \includegraphics[width=0.8\columnwidth]{log}
+        \includegraphics[width=0.9\columnwidth]{log}
         \caption{Screenshot of \pyRobots{} log analyser, processing the log of the
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \section{Detailed Example}
-% 
-% \begin{listing}[h!]
-% 
-% \begin{pythoncode}
-%     import time
-%     from robots import GenericRobot
-%     from robots.decorators import action, lock
-%     from robots.resources import Resource
-%     from robots.signals import ActionCancelled
-% 
-%     # create a 'lockable' resource for our robot
-%     WHEELS = Resource("wheels")
-% 
-%     class MyRobot(GenericRobot):
-% 
-%       def __init__(self):
-%         super(MyRobot, self).__init__()
-% 
-%         # create (and set) one element in the robot's state.
-%         # Here a sonar.
-%         self.state["sonar"] = 10.
-% 
-%         # do whatever other initialization you need :-)
-% 
-%       def send_goal(self, pose):
-%         # move your robot using your favorite middleware
-%         print("Starting to move towards %s" % pose)
-% 
-%       def stop(self):
-%         # stop your robot using your favorite middleware
-%         print("Motion stopped")
-% 
-%       def whatever_lowlevel_method_you_need(self):
-%         pass
-% 
-%     @lock(WHEELS)
-%     @action
-%     def move_forward(robot):
-%       """ Actions are written in a simple imperative, 
-%           blocking style.
-%       """
-% 
-%       # the target pose: simply x += 1.0m in the robot's 
-%       # frame. pyRobots will handle the frames 
-%       # transformations as needed.
-%       target = [1.0, 0., 0., "base_link"]
-% 
-%       try:
-%         robot.send_goal(target)
-% 
-%         while(robot.pose.distance(robot.pose.myself(), 
-%                                   target) > 0.1):
-%             # robot.sleep is exactly like time.sleep, 
-%             # except it lets the pyrobots signals pass 
-%             # through.
-%             robot.sleep(0.5)
-% 
-%         print("Motion succeeded")
-% 
-%       except ActionCancelled:
-%         # if the action is cancelled, clean up your state
-%         robot.stop()
-% 
-% 
-%     with MyRobot() as robot:
-% 
-%       # Turn on DEBUG logging.
-%       # Shortcut to configure the std Python logging system
-%       robot.debug()
-% 
-%       robot.whenever("sonar", below = 0.3).do(move_forward)
-% 
-%       try:
-%         while True:
-%           time.sleep(0.5)
-%       except KeyboardInterrupt:
-%         pass
-% \end{pythoncode}
-% \caption{Full example of a \pyRobots{} application.}
-% \label{mwe}
-% \end{listing}
-%
 
@@
         \toprule
-        Study                 & Robot & Middleware \\ \midrule
+        Study                 & Robot & Middlewares \\ \midrule
         Roboscopie            & PR2       & ROS, {\sc GenoM} \\
@@
 accessed from~\url{http://www.laas.fr/roboscopie}. Note that this script does
-not reflect anymore the toolkit usage.}
+not reflect anymore the current toolkit usage.}
 
@@
 \end{center}
-\caption{Examples of \pyRobots{} actions (extracted from experiments involving
-the PR2 robot).}
+\caption{Examples of \pyRobots{} actions.}
 
@@
 
+\begin{figure}
+        \centering
+        \includegraphics[width=0.9\columnwidth]{cowriter}
+        \caption{Children teaching an Aldebaran Nao interacting how to
+        handwrite.}
+        \label{expe-cowriter}
+\end{figure}
+
 \subsubsection{CoWriter} The {\sc CoWriter} project~\cite{hood2015those}
@@
 tools.
-
-\begin{figure}
-        \centering
-        \includegraphics[width=0.9\columnwidth]{cowriter}
-        \caption{Children teaching an Aldebaran Nao interacting how to
-        handwrite.}
-        \label{expe-cowriter}
-\end{figure}
 

### 2014-09-23 paper.tex
Implemented Ayberk's review. Thanks a lot!

@@
 
-Presented is \pyRobots{}, a new software toolkit for the executive control of
-complex robotic systems. Borrowing ideas from previous tools like {\sc
-urbi}~\cite{baillie2005urbi}, it proposes a lightweight Python framework to
+Presented is \pyRobots{}, a new open-source software toolkit for the executive
+control of complex robotic systems. Borrowing ideas from previous tools like
+{\sc urbi}~\cite{baillie2005urbi}, it proposes a lightweight Python framework to
 develop (pre-emptive) concurrent and event-based executive controllers.
@@
 unobstrusive tool suitable for rapid prototyping of complex interaction
-scenarios, \pyRobots{} also exposes several simple yet convenient abstractions for
-physical resources management and pose representation.
+scenarios, \pyRobots{} also exposes several simple yet convenient abstractions
+for physical resources management and pose representation. While
+middleware-agnostic, it features specific integration with the ROS middleware.
 
-Finally, and while middleware-agnostic, it features advanced integration with
-the \textit{de-facto} standard ROS middleware.
+Experimental deployments and stress-tests on several robotic platforms
+(including PR2 and Nao) in dynamic human environments are reported.
 
@@
 
-While individual decision-making components (like task planners) are studied in
-(relative) isolation since years, the \emph{orchestration} issue, with questions
-like \textit{When to start them? Which one should be selected? How to react to a
-new situation in a timely manner?} remains difficult to address in a generic
-way.
+While individual decision-making components (such as task planners) have been
+studied in (relative) isolation for decades, the \emph{orchestration} issue,
+with questions such as \textit{When to start them? Which one should be selected?
+How to react to a new situation in a timely manner?} remains difficult to
+address in a generic way.
 
-Many approaches have been devised (we review them in the next section) that
+Many approaches have been devised that
 include \emph{Finite State Machines} (FSM), domain-specific languages (in
-particular, logic languages) or agent-based frameworks.
+particular, logic languages) or agent-based frameworks; these are reviewed in
+the next section.
 
@@
 for a single application/demonstration/experiment and neither reliable nor
-extensible.
-
-We hypothesize that the software architecture design that these tools enforce,
-combined with their general lack of practicality (unfamiliar language; lack of
-bindings for a given robot or middleware; difficulty to install or setup;
-non-trivial deployment) explain this low level of acceptance more than any
-particular intrinsic weaknesses.
+extensible. We hypothesize that the software architecture that these tools
+enforce, combined with their general lack of practicality (unfamiliar language;
+lack of bindings for a given robot or middleware; difficulty to install or
+setup; non-trivial deployment) explain this low level of acceptance more than
+any particular intrinsic weaknesses.
 
 This article introduces our attempt at designing an unobtrusive execution
-control toolkit, that aims at addressing the \emph{practicality} issue: it has
+control toolset that aims at addressing the \emph{practicality} issue: it has
 been designed and implemented from the bottom-up, starting from actual needs
-when running large robotic systems (including Willow Garage's PR2 and
-Aldebaran's Nao) in dynamic and partially unpredictable environments (typically,
-loosely constrained human-robot interaction scenarios).
+when running large robotic systems (including robots such as PR2 or Nao) in
+dynamic and partially unpredictable environments, such as loosely constrained
+human-robot interaction scenarios.
 
@@
 rigid design and development guidelines, \pyRobots{} can be conceived as a set
-of software helpers to write parallel, event-based, high-level robot
+of software libraries to write parallel, event-based, high-level robot
 controllers.  It is written for and in Python, which ensure familiarity, fast
@@
 provides no mechanisms to connect to other modules or components. It instead
-\emph{relies} on one (or several) middleware to actually communicate with the
-robot and perform actions.
-
-Also, \pyRobots{} does not provide formal models or verifiability: in its
-current state, it does not attempt to contribute in this field. It focuses
-instead on providing a lightweight tool to control robots in environments that
-require carrying out many tasks in parallel, in an unpredictable, event-based
-manner. As such, \pyRobots{} is focusing on use-cases ranging from prototyping
-to mid-size experiments.
+\emph{relies} on one (or several) middleware(s) to actually communicate with the
+robot.
 
@@
 behaviours, excluding complete control stacks (that would typically include a
-middleware and hardware abstractions), meta-tools (like component generators or
-model checkers) and (semi-)automatic techniques (like emerging behaviours or
-learning by demonstration).  In contrast with middlewares for which the robotic
-community relies on a few \emph{de-facto} standards (ROS, YARP, OpenRTM), we
-want to underline that no such widely-used softwares exist for execution
-control, and we believe that most of today's robot behaviours are implemented as
-\emph{ad-hoc} scripts.
+middleware and hardware abstractions), meta-tools (such as component generators
+or model checkers) and (semi-)automatic techniques (such as emerging behaviours
+or learning by demonstration). Unlike \emph{middlewares} where a few tools --
+ROS, YARP, {\sc orocos}-rtt, OpenRTM -- are the \emph{de facto} standards, no tool really
+dominate the field of execution control, and we believe instead that most of
+today's robot behaviours are implemented as \emph{ad-hoc} scripts.
 
@@
 machines}, \emph{agent-based} controllers and
-\emph{domain-specific languages} (DSL), including \emph{embedded domain-specific
-languages} (eDSL,~\cite{joyeux2011robot}). These categories overlap to a certain
-extend, but they should reasonably reflect the motivational concepts/ideas that
+\emph{Domain-Specific Languages} (DSL), including \emph{embedded Domain-Specific
+Languages} (eDSL,~\cite{joyeux2011robot}). These categories overlap to a certain
+extent, but still reflect reasonably well the motivational concepts/ideas that
 gave birth to the specific tools.
 
-\paragraph{Finite State Machines} (or FSM) are the most well-known paradigm to
+\paragraph{Finite State Machines} They are the most well-known paradigm to
 describe how a robot goes from one \emph{state} (typically, performing an
-action) to another state when some condition becomes true or some event is
+action) to another state when some condition is satisfied or some event is
 triggered. Common FSM implementations in robotics include {\sc
-rfSM}~\cite{klotzbucher2010orocos} (part of the OROCOS framework) and
+rfSM}~\cite{klotzbucher2010orocos} (part of the {\sc orocos} framework) and
 {\sc smach}~\cite{bohren2010smach}. State machine semantics are well
 understood~\cite{harel1996statemate} and provide a well structured model,
-suitable for meta-reasoning (like verification). However, FSM are also
+suitable for meta-reasoning (such as verification). However, FSM are also
 acknowledged to be ill-suited for complex systems (where the number of state
 easily ``explodes'') or relatively dynamic and unpredictable environments where
-it can prove difficult to \textit{a priori} exhaustively list transitions.
+it can prove difficult to exhaustively list transitions \textit{a priori}.
 
@@
 to maintain the consistency of the global plan. ROAR~\cite{degroote2011roar}
-augments this idea with a logic mechanism allowing the system to automatically
+augments this idea with a logic mechanism, allowing the system to automatically
 select and start agents depending on the task at hand.
@@
 
-Domain-specific languages (DSL), in the broadest sense, represent the largest
-body of literature on high-level behaviour programming. They can either be
-actual domain-specific languages like PRS~\cite{ingrand1996prs},
-RPL~\cite{mcdermott1993reactive}, Golog~\cite{levesque1997golog} (and its
-derivatives like {\sc ReadyLOG}~\cite{ferrein2007robot}) or, more recently, {\sc
+In the broadest sense, they represent the largest body of literature on
+high-level behaviour programming. They can either be actual domain-specific
+languages such as PRS~\cite{ingrand1996prs}, RPL~\cite{mcdermott1993reactive},
+Golog~\cite{levesque1997golog} (and its derivatives such as {\sc
+ReadyLOG}~\cite{ferrein2007robot}) or, more recently, {\sc
 urbi}~\cite{baillie2005urbi}; or what Joyeux calls \emph{embedded}
-DSL~\cite{joyeux2011robot}: extensions of existing languages (typically, dynamic
-languages like Python or Ruby, or logic languages like Prolog) to provide
+DSL~\cite{joyeux2011robot}. eDSL are extensions of existing languages (typically, dynamic
+languages such as Python or Ruby, or logic languages such as Prolog) to provide
 high-level paradigms and control structures well suited to the programming of
@@
 similar fashion to {\sc roby} or {\sc teer}), that focuses on high-level
-behaviours implementation. It seeks unobtrusive interaction with the other
-components of the robot software stack, using Python as ubiquitous binding
-language. Similar to {\sc urbi}, it focuses on transparent pre-emptive
+behaviour implementation. Unlike larger stacks, \pyRobots{} does not provide
+formal models or verifiability. It aims to provide unobtrusive interaction with
+the other components of the robot software stack, using Python as the ubiquitous
+binding language. Similar to {\sc urbi}, it focuses on transparent pre-emptive
 concurrency and event-based programming.  Also, \pyRobots{}' model favours
 hierarchical task composition over plans, and does not provide by itself any
-task planning service, delegating this to optional external tools.
+task planning service.
 
@@
 
-This section introduces the main concepts of \pyRobots{},
-and provide an overview of the features and mechanisms. Since \pyRobots{} is an
-open-source software, the interested reader can also refer to the source code
-and documentation, available from~\url{https://github.com/chili-epfl/pyrobots}.
+This section introduces the main concepts of \pyRobots{}, and provide an
+overview of the features and mechanisms. The interested reader can refer to
+the source code and documentation, available
+from~\url{https://github.com/chili-epfl/pyrobots}.
 
@@
 \pyRobots{} core concepts are simple: on one hand, the user creates a
-\textbf{robot} as an instance of a \python|GenericRobot|, that encapsulates the
+\textbf{robot} as an instance of a \python|GenericRobot| that encapsulates the
 different low-level \textbf{controllers} (proxies to actuators, typically
-provided by a middleware like ROS) and the \textbf{state} of the robot
-(typically a set proxies to the sensors, or a connector to a database, etc.,
-accessed through a key-value container -- a dictionary-like object).
+provided by a middleware such as ROS) and the \textbf{state} of the robot
+(typically a set of proxies to the sensors accessed through a key-value
+container).
 
@@
 The user can list \textbf{events} that the \textbf{robot} must monitor, and
-attach callback to them.
+attach callbacks to them.
 
 Finally, the user can declare arbitrary \textbf{resources}, usually
-corresponding to the robot physical components (for instance, the wheels, the
-camera) and guarantee exclusive access to the required resource to specific
+corresponding to the robot's physical components (for instance, the wheels, the
+camera) and guarantee exclusive access to these resource to a specific set of
 actions.
@@
 independent threads. To the developer, however, they appear as regular Python
-function annotated with the \python|@action| decorator, and are invoked
+functions annotated with the \python|@action| decorator, and are invoked
 like any other function.  Since they are asynchronous, these functions may
-perfectly block or even never return (typically useful to implement continuous
+block or even never return (typically useful for implementing continuous
 background behaviours).
@@
 Actions can also call each other (nested actions), allowing natural
-decomposition of complex tasks into sub-actions and atomic actions (we note
-hereafter $depth$ the nesting level of an action, \ie the maximum number of
-calls to sub-actions until an atomic action is reached).
+decomposition of complex tasks into sub-actions (the nesting depth of an action
+is hereafter noted $depth$).
 
-\emph{Signaling:} Behaviours often need to be paused or cancelled.  This raises
+Behaviours often need to be paused or cancelled.  This raises
 an issue that specifically needs to be addressed in robotics: before suspending
@@
 middle of a step without first bringing back the foot to the ground). \pyRobots{}
-addresses this issue through signals, and extends standard Python thread into
+addresses this issue through signals, and extends the standard Python thread into
 \emph{signaling threads}, relying on Python's usual syntax for exception
 handling. Listing~\ref{lst:signals} illustrates this mechanism. The action is
-instantiated and started at line 8, and cancelled at line 10. This send a
+instantiated and started at line 8, and cancelled at line 10. This sends an
 \python|ActionCancelled| signal to the action that the user can handle using an
@@
 \python|safe_walk| action is cancelled. This sends the signal
-\python|ActionCancelled| to the action, that can appropriately handle it inside
+\python|ActionCancelled| to the action that can appropriately handle it inside
 the \python|except| block.}
@@
 
-Pausing actions is more complex (the action internal state must be saved, a mean
-to resume the action where it was interrupted must be provided, the user must be
-given ways to code transition behaviours when pausing and resuming) and is not
-currently supported by \pyRobots{}.
+Pausing actions is more complex (the action's internal state must be saved, a
+means to resume the action must be provided and the user must be given ways to
+code transition behaviours when pausing and resuming) and is not currently
+supported by \pyRobots{}.
 
@@
 through the methods \python|on| (one shot event) and \python|whenever|
-(continuous event monitoring). Event conditions can be either a predicate (that
-must take as single argument the robot instance) or a pair of $\{\text{state
-key}, \text{condition}\}$ with condition in $\{\text{value}=x, \text{become}=x,
+(continuous event monitoring). Event conditions can either be a predicate
+(taking the robot instance as its only argument) or a $\{\text{state
+key}, \text{condition}\}$ pair, with condition one of $\{\text{value}=x,
 \text{below}=x, \text{above}=x, \text{decrease}=x, \text{increase}=x\}$.
@@
 
-\subsection{Resources management}
+\subsection{Resource management}
 
 Physical robot resources (mainly actuators, but possibly sensors as
-well) usually mandate exclusive use (it is seldom desirable that two different actions
-control the wheels of a robot at the same time). \pyRobots{} provides a resource
+well) usually mandate exclusive use. \pyRobots{} provides a resource
 locking mechanism that relies on the \python|@lock| annotation: an action
 annotated with \python|@lock(ARMS)| (Listing~\ref{lst:resources}, l.6) will
-acquire (or wait for) the \python|ARM| resource when being executed.
+acquire (or wait for) the \python|ARMS| resource before being executed.
 
 This mechanism is generic and \pyRobots{} makes no assumptions concerning the
-actual resources available/meaningful to a specific robot. The
+actual resources available to a specific robot. The
 programmer define them himself, either as stand-alone or as compound
-resources (listing~\ref{lst:resources}, l.1--3).
+resources (Listing~\ref{lst:resources}, l.1--3).
 
-A given action can lock several distinct resources (with the corresponding
+A given action can lock several distinct resources with the corresponding
 number of \python|@lock| annotations. When executed, the
-action acquires a lock on the resource (and all dependent resources in the case
+action acquires a lock on the resource (and all sub-resources in the case
 of a compound resource) which is released upon action completion. If the lock is
 already acquired, two strategies are possible: either wait or \emph{fail fast}
-(\python|@lock(..., wait=False)|, l.16).
+(\python|@lock(..., wait=False)|, Listing~\ref{lst:resources}, l.16).
 
@@
 
-Managing poses, transformations, frames in real-world scenarios is not an easy
-task, mainly because it seems that every single library, middleware or robotic
-platform relies on its own conventions (rotation representations, in particular,
-are a common source of inconsistencies -- quaternions, (the many) Euler
+Managing poses, transformations and frames in real-world scenarios is not an
+easy task, mainly because it appears that libraries, middlewares or robotic
+platforms tend each to rely on their own conventions; rotation representations,
+in particular, are a common source of inconsistencies -- quaternions, Euler
 conventions, rotation matrices, Rodrigues, yaw-pitch-roll notations being all
-rather frequent). Within the ROS ecosystem, the Transform Library ({\sc
-tf},~\cite{foote2013tf}) has played a major role to make it convenient to work
-with frames, and poses can always be expressed and shared in the most convenient
+rather frequent. Within the ROS ecosystem, the Transform Library ({\sc
+tf},~\cite{foote2013tf}) has played a major role in making it convenient to work
+with frames: poses can always be expressed and shared in the most convenient
 local frame, {\sc tf} taking care of required intermediate transformation.
@@
 \pyRobots{} pose management builds on this idea to offer flexible pose
-representations \emph{within} the control scripts. Everytime a 6D pose needs to
+representations \emph{within} the control scripts. Every time a 6D pose needs to
 be specified (like a navigation goal or a target pose for an arm), it can be
-expressed as a simple frame name, by a triple $\{x, y, z\}$ (default frame {\tt
-map} is then assumed), by a 4-tuple $\{x, y, z, frame\}$, by a 6-tuple $\{x, y,
-z, rx, ry, rz\}$ (using $sxyz$ Euler convention), by a 7-tuple $\{x, y, z, rx,
-ry, rz, frame\}$, by a 7-tuple $\{x, y, z, qx, qy, qz, qw\}$, by a 8-tuple $\{x,
-y, z, qx, qy, qz, qw, frame\}$ or by a dictionary that defines any subset of
-these 8 values.  Internally, \pyRobots{} normalizes and stores the poses in
-their full 8-tuple expansion, as illustrated in listing~\ref{lst:poses}.
+expressed as a simple frame name, by a 3-tuple $(x, y, z)$ (default frame {\tt
+map} is then assumed), by a 4-tuple $(x, y, z, frame)$, by a 6-tuple $(x, y, z,
+rx, ry, rz)$ (using $sxyz$ Euler convention), by a 7-tuple $(x, y, z, rx, ry,
+rz, frame)$, by a 7-tuple $(x, y, z, qx, qy, qz, qw)$, by an 8-tuple $(x, y, z,
+qx, qy, qz, qw, frame)$ or by a dictionary that defines any subset of $(x, y, z,
+qx, qy, qz, qw, frame)$.  Internally, \pyRobots{} normalizes and stores the
+poses in their full 8-tuple expansion, as illustrated in
+Listing~\ref{lst:poses}.
 
@@
 
-Using helper methods like \python|inframe(<pose>, <other frame>)| or
-\python|pantilt(<pose>, <ref frame>)|, the programmer can then
+Using helper methods such as \python|inframe| \python|(<pose>, <other frame>)| or
+\python|pantilt| \python|(<pose>, <ref frame>)|, the programmer can then
 transform a given pose into a different frame or different representation.
@@
 Required transformations are computed by \pyRobots{}, and the programmer can
-provide new ``sources'' of frame transformations through the instanciation of
+provide new ``sources'' of frame transformations through the instantiation of
 \python|FrameProviders| that takes advantage of robot- and middleware-specific
@@
 
-Running robots in real-world, \ie dynamic and unpredictable, environments
+Running robots in real-world  environments that are dynamic and unpredictable
 entails repetitive coding-testing-debugging development cycles. \pyRobots{}
-provides several tools to support this development phase. A special attention
-was paid to the logging infrastructure. It relies on the versatile Python
-built-in logging tools, which support configurable parallel logging streams
-(typically, a full debug stream is logged to a file while a less verbose stream
-is output on the console at run-time).
+provides several tools to support this development phase that rely on the
+versatile Python built-in logging tools, which support configurable parallel
+logging streams (typically, a full debug stream is logged to a file while a less
+verbose stream is output on the console at run-time).
 
-\pyRobots{} logs are designed to allow full \emph{post-mortem} replay, and we
+\pyRobots{} logs are designed to allow full \emph{offline} replay, and we
 provide a graphical interface to display the actions and resource usage over
@@
         \includegraphics[width=0.9\columnwidth]{log}
-        \caption{Screenshot of \pyRobots{} log analyser, processing the log of the
-        experiment presented at section~\ref{croq}}
+        \caption{Screenshot of \pyRobots{} log analyzer, processing the log of the
+        experiment presented in section~\ref{croq}}
         \label{log_view}
@@
 Debugging and introspection support is provided at run-time through a range of
-off-the-shell Python debuggers. Remote debugging is also supported with standard
-tools like {\sc rpdb2}~\footnote{\url{http://winpdb.org}}. Thanks to the dynamic
-nature of Python, these debuggers not only provide insights on failures, but
-also provide full introspection and run-time manipulation of the application
-state.
+off-the-shelf Python debuggers. Remote debugging is also supported with standard
+tools such as {\sc rpdb2}~\footnote{\url{http://winpdb.org}}. Python being an
+interpreted language, these debuggers do not only provide insights on failures,
+but also tooling for full introspection and run-time manipulation of the
+application state.
 
@@
 18min long theatre performance between a human actor and a robot. Since that
-time, the toolkit has iteratively matured, and has been deployed and used on
-several projects and platforms. We first briefly recap these deployments with
-references to the corresponding publications, and then present our latest study,
-in a kindergarten, that recently acted as a stress-test for the library.
+time, the toolset has iteratively matured, and has been deployed and used on
+several projects and platforms. We first briefly summarize these deployments,
+and then present a recent study in a kindergarten that acted as a stress-test
+for the library.
 
@@
     \end{tabular}
-    \caption{Overview of significant experiments and studies where \pyRobots{}
-    has been used.}
+    \caption{Overview of the main studies where \pyRobots{} was used.}
 \end{table}
@@
 a theatre performance entitled \emph{Roboscopie}~\cite{lemaignan2012roboscopie}.
-The toolkit was created out of the need to control a complex robot (a Willow
-Garage PR2) in a structured way, as required for a live performance with a
+The toolset was created out of the need to control a complex robot (a PR2) in a structured way, as required for a live performance with a
 human\footnote{The original \pyRobots{} script used for the performance can be
 accessed from~\url{http://www.laas.fr/roboscopie}. Note that this script does
-not reflect anymore the current toolkit usage.}
+not reflect anymore the current toolkit usage.}.
 
 A total of 47 \pyRobots{} actions were written for this event
-(Table~\ref{pyrobots_actions} gives a few significant examples), and the toolkit
-proved to be especially convenient to interleave calls to both ROS-based and
+(Table~\ref{pyrobots_actions} gives a few significant examples), and the toolset
+proved to be convenient for interleaving calls to both ROS-based and
 {\sc GenoM}-based~\cite{mallet2010genom3} components through their respective
@@
 
-\pyRobots{} also allowed us to retain the \emph{rapid prototyping} convenience of a
-script language, essential for the robot to ``practise'' and adapt to the
-vision of the director.
+\pyRobots{} also allowed us to retain the \emph{rapid prototyping} convenience
+of a script language, essential for the robot's programmer to quickly adapt to
+the vision of the director.
 
@@
 is one of the two available executive controllers, and integrates with a task
-planner, a symbolic knowledge base, several motion planner and a natural
+planner, a symbolic knowledge base, several motion planners and a natural
 language processor. \pyRobots{} was favoured over the second execution
@@
         \includegraphics[width=0.9\columnwidth]{cowriter}
-        \caption{Children teaching an Aldebaran Nao interacting how to
-        handwrite.}
+        \caption{\textbf{CoWriter}: Children interactively teaching a Nao how to
+        write.}
         \label{expe-cowriter}
@@
 \subsubsection{CoWriter} The {\sc CoWriter} project~\cite{hood2015those}
-(Figure~\ref{expe-cowriter}) involves an Aldebaran Nao robot and young students
+(Figure~\ref{expe-cowriter}) involves a Nao robot and young students
 in a \emph{learning by teaching} interaction: the children have to teach the
-robot how to handwrite. This experiment involved a dual ROS/NAOqi environment,
+robot how to write. This experiment involved a dual ROS/NAOqi environment,
 bridged within \pyRobots{}. The high-level behaviour was implemented as a finite
-state machine ({\sc SMACH}) whose state corresponded to \pyRobots{} actions.
+state machine ({\sc smach}) whose states corresponded to \pyRobots{} actions.
 While the execution controller was subsequently simplified (towards a tighter
@@
 
-We organized in June 2014 a pilot study in a kindergarten (on campus) to
-stress-test the hardware and software of the Ranger
-robot~\cite{mondada2014ranger}. The Ranger robot is an autonomous robot designed
-for interaction with children. As it was primarily designed to support children
-with tidying up, it is shaped as a wooden ``box on wheels''
-(Figure~\ref{expe-nursery}). Hardware-wise, it features an RGB-D camera, IR
-sensors, a range and bearing sensor, a physical front bumper, a scale, a
-removable pacifier (used as a soft on/off switch by the children) and it is
-covered with capacitive tactile panels. The wooden panels hide 186 RGB LEDs,
-enabling various light patterns, it can play sounds, the two eyes can be fully
-controlled (2 DoFs pupils and actuated eye-lids). Wheel controllers are designed
-such as they switch to freewheel as soon as the robot is pushed by the children.
-The robot is powered by an embedded Gumstix Overo board (ARM7l) running a custom
-Linux image, assisted by three custom microcontroller boards (for low-level
-hardware interfacing).
+We organized a pilot study in a kindergarten to stress-test the hardware and
+software of the Ranger robot~\cite{mondada2014ranger}. The Ranger robot is an
+autonomous robot designed for interaction with children. As it was primarily
+designed to support children with tidying up, it is shaped as a wooden ``box on
+wheels'' (Figure~\ref{expe-nursery}). It features an RGB-D camera, IR sensors, a
+range and bearing sensor, a physical front bumper, a scale, a removable pacifier
+(used as a soft on/off switch by the children) and is covered with capacitive
+tactile panels. The wooden panels hide 186 RGB LEDs, enabling various light
+patterns, the two eyes can be fully controlled (2 DoFs pupils and actuated
+eye-lids) and the robot can play sounds. Wheel controllers are designed such
+that they switch to freewheel as soon as the robot is pushed by the children.
+The robot is powered by a Gumstix Overo board (ARM7l) running a custom Linux
+image, assisted by three custom microcontroller boards for low-level hardware
+interfacing.
 
-Software-wise, all the user-facing behaviours are implemented with \pyRobots{},
+All user-facing behaviours are implemented with \pyRobots{},
 and the communication with the hardware is abstracted through a mix of ROS
@@
         \includegraphics[width=0.9\columnwidth]{ranger-side}
-        \caption{Infants playing freely (!) with an EPFL's \emph{Ranger} robot.
+        \caption{\textbf{Stress-test}: Infants playing freely (!) with an EPFL's \emph{Ranger} robot.
         The behaviours of the robot are implemented via \pyRobots{}.}
@@
 We brought two such Ranger robots to a kindergarten (17 children, age: M=3.88,
-SD=0.65), and after a brief introduction (showing them that removing the
-pacifier would ``wake up'' the robot, and putting it back would make it ``fall
-asleep''), we invited the children to freely play with the robots and integrate
-them into their usual games. The experiment lasted 68 minutes.
+SD=0.65). After a brief introduction (showing them that removing the pacifier
+would ``wake up'' the robot, and putting it back would make it ``fall asleep''),
+we invited the children to freely play with the robots and integrate them into
+their usual games. The experiment lasted 68 minutes.
 
@@
 \caption{Simplified source of the high-level behaviours running on the robots during the
-kindergarten study (some behaviours like battery management have been omitted for
+kindergarten study (some behaviours such as the battery management have been omitted for
 clarity).}
@@
 
-Figure~\ref{croq_actions} shows the average number of events fired and action
-(including sub-actions) \emph{started} per minute, for the whole experiment. The
+Figure~\ref{croq_actions} shows the average number of events fired and actions
+(including sub-actions) started per minute for the whole experiment. The
 diagram reveals several peaks above 180 actions/minutes with a global average of
@@
 The results of this stress-test provide a reasonable baseline for the
-performances in terms of parallelism and debug/introspection capabilities that
-can be expected from an executive controller.
+performance in terms of parallelism and debug/introspection capabilities that
+can be expected from an executive controller in a human-robot ``open''
+interaction context.
 
@@
 
-We have presented here \pyRobots{}, a novel Python library supporting
+We have presented \pyRobots{}, a novel Python library supporting
 development of high-level executive controllers for robotics.
@@
 handling of task cancellation; support for event-based programming; a flexible
-robot resources locking mechanism and a built-in support for 6D pose
+robot resource locking mechanism and built-in support for 6D pose
 manipulation and transformation.
 
-\pyRobots{} does not aim at replacing large execution control frameworks. It
-proposes instead an accessible set of tools to improve upon the all-too-common
+\pyRobots{} does not aim to replace large execution control frameworks. Instead,
+it provides an accessible set of tools to improve upon the all-too-common
 \emph{ad-hoc} scripts, while remaining middleware agnostic.
 
-The article also reports on several experimental deployments on four different
-kind of robots using a mixture of four different middlewares.
+We also report on several experimental deployments on four different
+kind of robots using four different middlewares.
 
-While already four years old, several possible developments would offer
-interesting perspectives for this tool, including (a)~the support to suspend (and
-then resume) actions (which requires defining a satisfying syntax to declare
-suspend/resume behaviours within actions); (b)~mechanisms to define action
-priorities; (c)~investigate recent work on Python's {\tt asynchio} mechanisms
-(that provide reference implementation of co-routines in Python).
+Several possible developments would offer interesting perspectives for this
+tool, including (a)~the support to suspend (and then resume) actions which
+requires defining a satisfying syntax to declare suspend/resume behaviours
+within actions; (b)~mechanisms to define action priorities; (c)~investigate
+recent work on Python's {\tt asynchio} mechanisms (that provide reference
+implementation of co-routines in Python).
 
 Being available as an open-source platform, documented and easy to install, we
-believe that these development objectives may materialize soon, as well-defined
+believe that these development objectives may be reached soon, as well-defined
 use-cases emerge from the library's users.

### 2014-09-29 root.tex
paper.tex -> root.tex (mandated by ICRA)

--- 2014-09-23 paper.tex
@@
 ROS, YARP, {\sc orocos}-rtt, OpenRTM -- are the \emph{de facto} standards, no tool really
-dominate the field of execution control, and we believe instead that most of
+dominates the field of execution control, and we believe instead that most of
 today's robot behaviours are implemented as \emph{ad-hoc} scripts.

### 2015-02-27 root.tex
Adaptation to resubmit to IROS

@@
 %        1         2         3         4         5         6         7         8
-\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper
+\documentclass[letterpaper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper
 
@@
 
-\subsubsection{CoWriter} The {\sc CoWriter} project~\cite{hood2015those}
+\subsubsection{CoWriter} The {\sc CoWriter} project~\cite{hood2015when}
 (Figure~\ref{expe-cowriter}) involves a Nao robot and young students
